\section{Keller fibers and the rank theorem}

Let $K$ be a field of characteristic zero.  A polynomial map
\[
 F:\A^m_K\longrightarrow\A^m_K
\]
is a \emph{Keller map} if $\det DF\in K^*$.  Its geometric degree is
\[
 \gdeg(F)=[K(x_1,\ldots,x_m):K(F_1,\ldots,F_m)].
\]
A Keller map is \'etale and hence quasi-finite.  Every fiber over a $K$-point
is therefore a finite \'etale $K$-scheme, possibly empty.

\begin{definition}[Keller fiber]\label{def:keller-fiber}
A nonzero finite \'etale $K$-algebra $A$ is a \emph{Keller fiber} if there
are a Keller map $F:\A^m_K\to\A^m_K$ and a point $y\in\A^m(K)$ such that
\[
 F^{-1}(y)\simeq\Spec A,
 \qquad \dim_KA=\gdeg(F).
\]
Equivalently, $\Spec A$ is a full fiber: it contains every generic inverse
sheet, so no sheet is lost at the normalization boundary.  An arbitrary
special fiber can be shorter because inverse sheets may escape through the
nonproperness boundary; fullness excludes precisely this phenomenon.
\end{definition}

The construction can be summarized in one line:
\[
 \boxed{
 \text{squarefree }P(T)
 \ \longmapsto\
 G(S)=P(a+S)-P(a)
 \ \longmapsto\
 \widetilde F_G^{-1}(y_{P,a})\simeq\Spec K[T]/(P).}
\]

\begin{lemma}[The degree-two exclusion over $K$]\label{lem:degree-two}
Let $K$ be any field of characteristic zero.  A Keller map
$F:\A^m_K\to\A^m_K$ of geometric degree two is a polynomial automorphism.
\end{lemma}

\begin{proof}
Put $L=K(x_1,\ldots,x_m)$ and $M=K(F_1,\ldots,F_m)$.  The extension
$L/M$ has degree two, hence is separable and Galois.  The Galois case of the
Jacobian theorem was proved over $\mathbb C$ by Campbell and subsequently by
Razar and independently Wright in algebraic form
\cite{Campbell,Razar,Wright}.

For completeness, we spell out why this gives the exact ground-field
generality used here.  Choose a finitely generated subfield
$K_0\subset K$ containing all coefficients of $F$, and regard $F$ as a map
over $K_0$.  After shrinking the target, this map is finite locally free; its
rank is its generic degree.  Flat base change preserves that rank.  Since its
base change to $K$ has generic degree two, the map over $K_0$ also has generic
degree two, and so does every further scalar extension.

Choose an embedding $K_0\hookrightarrow\mathbb C$.  The complexified
function-field extension is therefore quadratic and Galois, so Campbell's
theorem makes $F_{\mathbb C}$ a polynomial automorphism.

The formal inverse of $F$ at the point $F(0)$ is unique.  Its coefficients
are obtained recursively from the coefficients of $F$ and the inverse of its
linear Jacobian, and hence lie in $K_0$.  The polynomial inverse over
$\mathbb C$ agrees with this formal inverse, so all of its coefficients lie
in $K_0$.  It therefore descends to $K$, proving the assertion without any
algebraic-closedness assumption.
\end{proof}

\begin{theorem}[Finite-\'etale realization]\label{thm:main}
Over a field of characteristic zero, the possible ranks of nonzero Keller
fibers are exactly
\[
 \boxed{1,3,4,5,\ldots}.
\]
More precisely, let $P\in K[T]$ be squarefree of degree $N\geq3$.  Choose
$a\in K$ outside the finite zero set of $P'P'''$, and put
\[
 G(S)=P(a+S)-P(a).
\]
Let $F_G=(\Pi,B,C)$ be the quadratic-gauge map of
\eqref{eq:gauge-map}, and define
\[
 \widetilde F_G=L\circ F_G,
 \qquad
 L(\Pi,B,C)=\left(\Pi,-\frac{B}{2},C\right).
\]
Then
\[
 \det D\widetilde F_G=1,
 \qquad \gdeg(\widetilde F_G)=N,
 \qquad \max_i\deg (\widetilde F_G)_i\leq 6N+2.
\]
At
\begin{equation}
 y_{P,a}=\left(1,0,-\frac{2P(a)}{P'(a)}\right)
 \label{eq:main-target}
\end{equation}
its full fiber is
\[
 \boxed{\widetilde F_G^{-1}(y_{P,a})
 \simeq\Spec K[T]/(P).}
\]
Consequently every finite \'etale $K$-algebra of rank at least three is a
Keller fiber.
\end{theorem}

\begin{proof}
The translation makes the linear and cubic coefficients of $G$ equal to
$P'(a)$ and $P'''(a)/6$, respectively, so the quadratic gauge of
Section~\ref{sec:gauge} applies.  At the target \eqref{eq:main-target}, its
inverse polynomial is
\[
 G(S)-\frac{P'(a)}2\left(-\frac{2P(a)}{P'(a)}\right)
 =P(a+S).
\]
Proposition~\ref{prop:gauge} identifies the full scheme fiber with
$\Spec K[S]/(P(a+S))$, which is $\Spec K[T]/(P)$ after $T=a+S$.
The linear automorphism $L$ fixes this target because its second coordinate
is zero, multiplies the Jacobian determinant by $-1/2$, and does not alter
geometric degree or coordinate degrees.  The determinant-one claim and the
effective bound follow from Corollary~\ref{cor:effective-gauge}.

For an arbitrary finite \'etale algebra of rank at least three, apply the
monogenicity lemma in Section~\ref{sec:realization}.  Rank one is realized by
the identity, and Lemma~\ref{lem:degree-two} excludes rank two.  Finally,
$P'P'''$ is nonzero for $N\geq3$ in characteristic zero, and $K$ is infinite,
so the required $a$ always exists.
\end{proof}
