\section{The smallest Hasse-principle failure}\label{sec:hasse}

For a number field $K$, call a Keller fiber $A$ a \emph{Hasse failure} if
$\Spec A$ has a point over every completion of $K$ but no $K$-rational
point.

We first isolate the elementary lower bound.

\begin{lemma}[Degree-four barrier]\label{lem:degree-four-barrier}
A nonempty finite \'etale $\Q$-scheme of degree at most four which has a
point over every $\Q_p$ has a rational point.
\end{lemma}

\begin{proof}
Let $X$ be such a scheme and let $L$ split it.  For every unramified prime,
$X(\Q_p)\neq\varnothing$ exactly when the Frobenius element fixes a point of
the finite $\operatorname{Gal}(L/\Q)$-set $X(L)$.

If $X$ is connected of degree greater than one, this action is transitive and
contains a derangement.  Indeed, Burnside's lemma says that the average
number of fixed points is one, while the identity fixes more than one point;
hence some element fixes none.  Chebotarev supplies an unramified prime with
that Frobenius class, contradicting local solubility.

It remains to consider disconnected schemes without a rational component.
In total degree at most four the only possibility is the disjoint union of
two quadratic fields.  If they coincide, any inert prime works.  If they are
distinct, their compositum has Galois group $C_2\times C_2$ and contains an
element nontrivial in both quadratic quotients.  Chebotarev again supplies a
prime inert in both components.  Thus an everywhere locally soluble scheme
of degree at most four must contain $\Spec\Q$.
\end{proof}

Berend and Bilu's classical example is
\begin{equation}
 P_5(T)=(T^3-19)(T^2+T+1).
 \label{eq:BB}
\end{equation}
They use it in their study of polynomials with roots modulo every integer
\cite{BerendBilu}.

\begin{theorem}[Optimal Hasse-failing Keller fiber]\label{thm:hasse}
There is an explicit Keller map $F:\A^3_\Q\to\A^3_\Q$ of geometric degree
five and a rational target $y$ such that
\[
 F^{-1}(y)(\Q)=\varnothing,
 \qquad
 F^{-1}(y)(\Q_v)\neq\varnothing
 \quad\text{for every place }v.
\]
No Hasse-failing Keller fiber has rank below five.  Hence
\[
 \boxed{d_{\mathrm{HP}}=5}.
\]
\end{theorem}

\begin{proof}
The polynomial \eqref{eq:BB} has no rational root.  For a prime $p\neq3$,
if $p\equiv1\pmod3$, then $T^2+T+1$ has a simple root modulo $p$; if
$p\equiv2\pmod3$, cubing is bijective on $\mathbb F_p^*$, so $T^3-19$ has a
simple root modulo $p$.  At $p=3$, write $T=1+3u$; the equation $T^3=19$
reduces after division by $9$ to
\[
 u+3u^2+3u^3=2,
\]
which has the simple solution $u\equiv2\pmod3$ and therefore lifts.  The
cubic factor has a real root.  Thus \eqref{eq:BB} has a root over every
completion of $\Q$.

The explicit map is given below.  Its fiber over the stated target is exactly
$\Spec\Q[T]/(P_5)$, so it has the same local and global points.
Lemma~\ref{lem:degree-four-barrier} proves minimality.
\end{proof}

We now derive the displayed map from the normalized gauge, including the
scaling rather than treating its determinant as a separate computation.  Put
\[
 G(T)=P_5(T)-P_5(0)=T^5+T^4+T^3-19T^2-19T.
\]
Thus
\[
 (g_1,g_2,g_3,g_4,g_5)=(-19,-19,1,1,1).
\]
With $t=1+xy$ and
\[
 q=t^2z-19y^2(1+3t),
\]
Proposition~\ref{prop:gauge} gives the normalized map
\begin{align*}
 \Pi={}&tq,\\
 B_0={}&y-\frac3{19}xq+2tq
       -\frac4{19}t^2x^2q^4-\frac5{19}t^2x^3q^5,\\
 C_0={}&x(5-3t)+\frac1{19}x^3z
       +\frac2{19}(xq)^4+\frac3{19}(xq)^5,
\end{align*}
with $\det D(\Pi,B_0,C_0)=-2$.  Apply the target linear automorphism
\[
 L(\Pi,B_0,C_0)=(\Pi,19B_0,19C_0).
\]
The resulting denominator-free map $F=L\circ F_G$ is
\begin{equation}
\begin{aligned}
 F_1={}&tq,\\
 F_2={}&19y-3xq+38tq-4t^2x^2q^4-5t^2x^3q^5,\\
 F_3={}&19x(5-3t)+x^3z+2(xq)^4+3(xq)^5.
\end{aligned}
\label{eq:explicit-hasse-map}
\end{equation}
Consequently
\[
 \boxed{\det DF=\det(L)(-2)=19^2(-2)=-722.}
\]
The normalized target $(1,0,-2)$ is carried by $L$ to
\[
 y=(1,0,-38).
\]
At the normalized target, the inverse equation is
\[
 G(T)-\frac{-19}{2}(-2)=G(T)-19=P_5(T).
\]
Proposition~\ref{prop:gauge} therefore gives the scheme-theoretic identity
\begin{equation}
 \boxed{F^{-1}(1,0,-38)\simeq
 \Spec\Q[T]/\bigl((T^3-19)(T^2+T+1)\bigr).}
 \label{eq:explicit-hasse-fiber}
\end{equation}
Equations \eqref{eq:explicit-hasse-map}--\eqref{eq:explicit-hasse-fiber} are
an exact certificate of the smallest example.
