\section{Consequences, provenance, and literature boundary}

Theorem~\ref{thm:main} is a realization theorem, not merely a source of
counterexamples.  Some immediate applications are:
\begin{enumerate}
\item Every squarefree polynomial of degree at least three is, after one
      translation of its variable, the exact inverse polynomial of a full
      Jacobian-one Keller fiber.
\item A degree-$N$ input has an explicit realization in $\A^3$ of coordinate
      degree at most $6N+2$.
\item Every intersective polynomial of degree at least three transfers to a
      Hasse-failing Keller fiber without stabilization or auxiliary factors.
\item Any real signature, splitting field, or collection of local
      factorization patterns already carried by a finite \'etale algebra is
      carried by the corresponding Keller fiber, compatibly with extension of
      the ground field.
\item Connected fibers require no additional geometric argument: start with
      an irreducible polynomial.
\end{enumerate}

The fourth item is deliberately a transfer statement.  It does not solve an
inverse Galois problem or assert that arbitrary independently prescribed
local algebras globalize; it says that every arithmetic realization which
already exists is preserved exactly by the Keller construction.

The realization is not unique.  Separate boundary calculations in the
accompanying research repository show that the same finite \'etale algebra
can occur in stably inequivalent weighted, cancellation, and quadratic-gauge
Keller maps.  That stable-classification problem is logically independent of
the present theorem and is best treated in a sequel, tentatively titled
\emph{One Fiber, Many Keller Maps}.

\paragraph{Literature boundary.}
A dated search completed on 24 July 2026 found no earlier statement that every
finite \'etale algebra of rank at least three is a full fiber of a polynomial
Keller map, nor the resulting rank classification.  This is necessarily a
qualified priority statement: bibliographic searches can miss unpublished
work, differently phrased results, and sources not indexed electronically.
The appropriate claim is therefore ``to our knowledge'', not an absolute
firstness assertion.  A reproducible search ledger is included with the
paper sources.

The closest results located have different scopes.  Gallagher gives explicit
Keller maps in every geometric degree and complete split rational fibers
\cite{Gallagher}; this realizes $K^N$, rather than an arbitrary finite
\'etale $K$-algebra.  Miranda--Neto proves radicality and maximal-ideal
decomposition properties for fibers already arising from a Keller map
\cite{Miranda}; this is not a prescribed-fiber realization theorem.
Lipton and Markakis study rational images and Hilbert irreducibility
\cite{LiptonMarkakis}, while Shaska studies graded Keller maps and arithmetic
thinness phenomena \cite{Shaska}.  None of these sources states the
finite-\'etale realization theorem proved here.

\paragraph{Exact verification.}
The companion checker
\[
 \texttt{scripts/verify\_finite\_etale\_keller\_fibers.py}
\]
verifies the source-chart and marked-line identities, two-sided
scheme-theoretic reconstruction in quotient rings, the fixed Jacobian-one
normalization, the effective coordinate-degree bound on independent translated
examples, the exact scaling of the explicit degree-five map, its determinant,
the inverse-polynomial identities, and the fixed-map infinite family.  The
universal proof is the calculation in Proposition~\ref{prop:gauge} and
Corollary~\ref{cor:effective-gauge}, not a bounded computer search.

\paragraph{Lean formalization.}
The repository contains a staged, axiom-free Lean project at
\[
 \texttt{formal/finite-etale-keller}.
\]
Its first stage certifies the explicit quintic Jacobian determinant, both
output normalizations, the inverse-polynomial identity, and a constructive
B\'ezout inverse for the derivative class in the quotient algebra.  The next
formal stages are the universal marked-line identities, the two-sided
coordinate-ring reconstruction, and the translated polynomial-realization
theorem.  The historical degree-two Galois case is kept separate from the new
constructive core until that classical input is itself formalized.

\paragraph{Provenance.}
The recent three-dimensional counterexample was publicly attributed to
Alp\"oge and Fable.  Gallagher independently developed a weighted all-degree
viewpoint and an exact public atlas of complete rational fibers
\cite{Gallagher}.  The proof here instead uses the root-engineered
quadratic-gauge construction and the finite-\'etale realization theorem.
Berend and Bilu are credited for the classical intersective quintic.  The
rank-two exclusion is the Galois case proved over $\mathbb C$ by Campbell and
subsequently treated algebraically by Razar and independently Wright
\cite{Campbell,Razar,Wright}; Lemma~\ref{lem:degree-two} includes the descent
argument needed for an arbitrary characteristic-zero ground field.

\paragraph{AI assistance.}
Generative AI tools were used for symbolic experimentation, exact checker
development, adversarial proof review, literature search, formalization, and
manuscript preparation.  The author reviewed the mathematical claims and
assumes full responsibility for the contents.

\begin{thebibliography}{11}
\small

\bibitem{BerendBilu}
D.~Berend and Y.~Bilu,
\emph{Polynomials with roots modulo every integer},
Proc. Amer. Math. Soc. \textbf{124} (1996), 1663--1671,
\href{https://doi.org/10.1090/S0002-9939-96-03210-8}
{doi:10.1090/S0002-9939-96-03210-8}.

\bibitem{Campbell}
L.~A. Campbell,
\emph{A condition for a polynomial map to be invertible},
Math. Ann. \textbf{205} (1973), 243--248,
\href{https://doi.org/10.1007/BF01349234}
{doi:10.1007/BF01349234}.

\bibitem{Gallagher}
A.~Gallagher,
\emph{Counterexample atlas: generic fiber degrees 3 through 100},
online exact certificate atlas, 2026,
\href{https://jacobianfun.org/counterexamples}{jacobianfun.org}.

\bibitem{LiptonMarkakis}
R.~J. Lipton and E.~Markakis,
\emph{Some remarks on the Jacobian conjecture and connections with
Hilbert's irreducibility theorem},
\href{https://arxiv.org/abs/math/0507525}{arXiv:math/0507525}, 2005.

\bibitem{Miranda}
C.~B. Miranda--Neto,
\emph{An ideal-theoretic approach to Keller maps},
Proc. Edinburgh Math. Soc. \textbf{62} (2019), 1033--1044,
\href{https://doi.org/10.1017/S0013091519000099}
{doi:10.1017/S0013091519000099}.

\bibitem{Razar}
M.~Razar,
\emph{Polynomial maps with constant Jacobian},
Israel J. Math. \textbf{32} (1979), 97--106,
\href{https://doi.org/10.1007/BF02764906}
{doi:10.1007/BF02764906}.

\bibitem{Shaska}
T.~Shaska,
\emph{Graded Keller maps and the Jacobian conjecture},
\href{https://arxiv.org/abs/2607.20210}{arXiv:2607.20210}, 2026.

\bibitem{Wright}
D.~Wright,
\emph{On the Jacobian conjecture},
Illinois J. Math. \textbf{25} (1981), 423--440,
\href{https://doi.org/10.1215/ijm/1256047158}
{doi:10.1215/ijm/1256047158}.

\end{thebibliography}
