\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,microtype}
\usepackage[T1]{fontenc}
\usepackage[hidelinks]{hyperref}
\usepackage{booktabs}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newcommand{\A}{\mathbb A}
\newcommand{\Q}{\mathbb Q}
\newcommand{\Gm}{\mathbb G_m}
\newcommand{\Spec}{\operatorname{Spec}}
\newcommand{\Norm}{\operatorname{Norm}}
\newcommand{\rank}{\operatorname{rank}}
\newcommand{\Fitt}{\operatorname{Fitt}}
\newcommand{\cR}{\mathcal R}
\numberwithin{equation}{section}

\title{Common Arithmetic Fibers of\\Stably Inequivalent Keller Maps}
\author{Roy van Rijn\\{\small Independent researcher}}
\date{24 July 2026}
\hypersetup{
  pdftitle={Common Arithmetic Fibers of Stably Inequivalent Keller Maps},
  pdfauthor={Roy van Rijn},
  pdfsubject={Arithmetic fibers and stable inequivalence of Keller maps},
  pdfkeywords={Keller map, finite etale algebra, complete fiber,
    stable left-right equivalence, arithmetic fiber}
}

\begin{document}
\maketitle

\begin{abstract}
We show that a complete regular arithmetic fiber does not determine its
ambient polynomial Keller map, even up to polynomial left-right equivalence
and identity stabilization.  For every \(N\geq4\), two fixed Keller maps
over \(\Q\), of geometric degree \(N\), share the complete fiber
\[
 \Spec\Q[T]/(T^N+T^3-2T^2+T+u)
\]
for every squarefree rational specialization \(u\), but their stable classes
are distinct.  A Hilbert subset of specializations gives connected fibers.
Over \(K=\Q(\sqrt{-2})\), three fixed pairwise stably inequivalent quartic
Keller maps share every squarefree member of one quartic pencil; again a
Hilbert subset is connected.  The separation is intrinsic: the normalized
ramified boundary strata have different unit ranks, and the two reciprocal
constructions are distinguished by a relative Fitting divisor, or
independently by boundary-contact thickness.  We give small irreducible
quartic examples over both fields and exact finite-field certificates.
\end{abstract}

\section{The question and the results}

A polynomial map \(F:\A^n_K\to\A^n_K\) over a characteristic-zero field is
a \emph{Keller map} if its Jacobian determinant is a nonzero constant.
We call a fiber complete if it is regular and its length equals the
geometric degree of \(F\).  Thus no generic inverse sheet is lost at the
boundary.

Two maps are polynomially left-right equivalent if
\[
 F_2=\beta\circ F_1\circ\alpha
\]
for polynomial automorphisms \(\alpha\) and \(\beta\) of source and target.
They are \emph{stably} equivalent if their products with identity maps on
affine spaces become left-right equivalent.

For a finite etale \(K\)-algebra \(A\), define
\begin{equation}
 \cR_K(A)=
 \left\{
 \begin{array}{c}
 \text{stable polynomial left-right classes of Keller maps}\\
 \text{having a complete fiber isomorphic to }\Spec A
 \end{array}
 \right\}.
 \label{eq:R-definition}
\end{equation}
The definition is unmarked: the target carrying the fiber may vary.
Completeness makes the condition invariant under left-right equivalence,
and stabilization replaces the fiber by its product with the origin.

Our first result holds in every degree.

\begin{theorem}[Fixed pairs over \(\Q\)]\label{thm:rational-pairs}
For every \(N\geq4\), there are two fixed \(\Q\)-defined Keller maps
\[
 F_N^{\rm wt},F_N^{\rm quad}:\A^3\longrightarrow\A^3
\]
of geometric degree \(N\) with the following properties.
\begin{enumerate}
 \item For every \(u\in\Q\) such that
 \[
  P_{N,u}(T)=T^N+T^3-2T^2+T+u
  \label{eq:rational-pencil}
 \]
 is squarefree, both maps have a complete fiber isomorphic to
 \(\Spec\Q[T]/(P_{N,u})\).
 \item The two maps are not polynomially left-right equivalent, even after
 arbitrary identity stabilization.
 \item The polynomial \(P_{N,U}\) is irreducible over \(\Q(U)\).
 Consequently a Hilbert subset of \(u\in\Q\) gives connected common fibers,
 and for each such \(u\),
 \[
  \left|\cR_{\Q}\bigl(\Q[T]/(P_{N,u})\bigr)\right|\geq2.
 \]
\end{enumerate}
\end{theorem}

The second result replaces a pair by a triple.

\begin{theorem}[Fixed triples over \(\Q(\sqrt{-2})\)]
\label{thm:quadratic-triples}
Let \(K=\Q(\eta)\), where \(\eta^2=-2\), and put
\begin{equation}
 \Psi_R(T)=T-\frac92T^2+(8+2\eta)T^3
 -\left(\frac72+2\eta\right)T^4-R.
 \label{eq:triple-pencil}
\end{equation}
There are three fixed \(K\)-defined Keller maps
\[
 F^{\rm wt},F^{\rm can},F^{\rm quad}:\A^3_K\longrightarrow\A^3_K
\]
of geometric degree four such that:
\begin{enumerate}
 \item for every \(R\in K\) for which \(\Psi_R\) is squarefree, all three
 maps have a complete fiber isomorphic to \(\Spec K[T]/(\Psi_R)\);
 \item the three maps are pairwise stably polynomially left-right
 inequivalent;
 \item \(\Psi_R\) is irreducible over \(K(R)\), so a Hilbert subset of
 \(R\in K\) gives connected common fibers and
 \[
  \left|\cR_K\bigl(K[T]/(\Psi_R)\bigr)\right|\geq3.
 \]
\end{enumerate}
\end{theorem}

The maps are explicit, but the common fibers are best understood through
their one-variable inverse equations.  We recall the two programmable
constructions next, apply them to the pencils
\eqref{eq:rational-pencil} and \eqref{eq:triple-pencil}, and finish with the
stable boundary invariants.

\section{Two programmable inverse constructions}

\subsection{The weighted tangent construction}

Let \(P\in K[T]\) be squarefree of degree at least three, and set
\[
 H(T)=P(T)-P(0)-P'(0)T.
\]
The polynomial \(P\) is \emph{tangent-admissible} if
\begin{equation}
 P(1)-P(0)=P'(0),\quad
 c:=P'(0)-P'(1)\neq0,\quad
 P''(1)\neq2(P'(1)-P'(0)).
 \label{eq:tangent-admissible}
\end{equation}
Only \(H\) and \(c\), not the constant term of \(P\), enter the map.

Write \(p=H'\), \(\kappa=H''(1)/c\), and
\[
 a=-\frac{1+\kappa}{2+\kappa},\qquad
 L_H(W)=\int_0^W\frac{V H''(V)}c\,dV.
\]
In source coordinates put
\[
 u=1+xy,\qquad \gamma=1+axy+x^2z,\qquad W=u\gamma.
\]
The weighted map is
\begin{equation}
 F_H^{\rm wt}(x,y,z)=
 \left(
 \frac{u+L_H(W)/\gamma^2}{x^2},
 \frac{c+p(W)/\gamma}{x},
 x\gamma
 \right).
 \label{eq:weighted-map}
\end{equation}
The endpoint identities in \eqref{eq:tangent-admissible} make the displayed
quotients polynomial.

\begin{proposition}[Weighted inverse]\label{prop:weighted}
The map \eqref{eq:weighted-map} satisfies
\[
 \det DF_H^{\rm wt}=c.
\]
Writing its target coordinates as \((A,B,C)\), its inverse equation is
\begin{equation}
 E_{A,B,C}(W)=H(W)-BCW+cAC^2.
 \label{eq:weighted-inverse}
\end{equation}
At
\[
 (A,B,C)=\left(\frac{P(0)}c,-P'(0),1\right),
\]
equation \eqref{eq:weighted-inverse} is exactly \(P(W)=0\).  If \(P\) is
squarefree, reconstruction identifies the complete scheme fiber with
\(\Spec K[W]/(P)\).
\end{proposition}

\begin{proof}
The weighted tangent identities are
\[
 BC=H'(W)+c\gamma,\qquad
 cAC^2=W(H'(W)+c\gamma)-H(W).
\]
Eliminating \(\gamma\) gives \eqref{eq:weighted-inverse}.  Its derivative
gives
\[
 \gamma=-E_{A,B,C}'(W)/c,\qquad x=-cC/E_{A,B,C}'(W).
\]
Thus every simple root reconstructs exactly one finite source point.
The same two identities reduce the three-variable Jacobian to the plane
tangent determinant and give \(\det DF_H^{\rm wt}=c\).
\end{proof}

For the stable comparison we need one additional open condition: \(H\) is
\emph{boundary-clean} when zero is an exact double root and every nonzero
root is simple.

\subsection{The root-engineered quadratic gauge}

Let
\[
 G(S)=g_1S+\cdots+g_NS^N,\qquad g_1g_3g_N\neq0.
\]
Put
\[
 t=1+xy,\qquad
 q=t^2z+\frac{g_1}{g_3}y^2(1+3t).
\]
The root-engineered quadratic-gauge map is
\begin{align}
 \Pi={}&tq,\notag\\
 B={}&y+3\frac{g_3}{g_1}xq+2\frac{g_2}{g_1}tq
       +\sum_{k=4}^N k\frac{g_k}{g_1}
          t^2x^{k-2}q^k,\notag\\
 C={}&x(5-3t)-\frac{g_3}{g_1}x^3z
       -\sum_{k=4}^N(k-2)\frac{g_k}{g_1}(xq)^k.
 \label{eq:quadratic-map}
\end{align}

\begin{proposition}[Quadratic-gauge inverse]\label{prop:quadratic}
The map \eqref{eq:quadratic-map} satisfies \(\det DF_G^{\rm quad}=-2\).
Define
\[
 G_\Pi(S)=g_1S+\Pi(g_2S^2+g_3S^3)
 +\sum_{k=4}^Ng_k\Pi^kS^k.
\]
Its inverse equation is
\begin{equation}
 G_\Pi(S)-\frac{g_1}{2}(BS^2+C)=0.
 \label{eq:quadratic-inverse}
\end{equation}
At \((\Pi,B,C)=(1,0,0)\) this is \(G(S)=0\).  More generally, changing
only \(C\) adds an arbitrary constant to the inverse polynomial.  Every
simple root reconstructs exactly one source point.
\end{proposition}

\par\medskip\noindent\textit{Proof.}\ 
On \(t\neq0\), use \(S=x/t\) and \(Q=y+xq\).  Then
\[
 D=1-SQ+\Pi S^2=t^{-1}.
\]
Differentiating \eqref{eq:quadratic-inverse} on the source incidence gives
\[
 \partial_SE=g_1D.
\]
The rational source chart contributes the reciprocal factor \(D^{-1}\);
the plane incidence contributes \(-2D\).  Their product gives determinant
\(-2\), and \(D\neq0\) gives unique reconstruction for every simple root.
\hfill\(\square\)

\section{Two fixed maps with infinitely many common fibers}

Fix \(N\geq4\), and set
\begin{equation}
 H_N(T)=T^N+T^3-2T^2,\qquad
 P_{N,u}(T)=H_N(T)+T+u.
 \label{eq:HN}
\end{equation}
The basic arithmetic identities are independent of \(u\):
\[
 P_{N,u}(1)-P_{N,u}(0)=P_{N,u}'(0)=1,
\]
\[
 P_{N,u}'(0)-P_{N,u}'(1)=1-N\neq0,
\]
\[
 P_{N,u}''(1)=N(N-1)+2\neq2(N-1).
\]
Thus every squarefree member is tangent-admissible, and all members use the
same weighted map \(F_{H_N}^{\rm wt}\).  Proposition~\ref{prop:weighted}
recovers \(P_{N,u}\) at
\begin{equation}
 q_u^{\rm wt}=\left(\frac{u}{1-N},-1,1\right).
 \label{eq:weighted-target-N}
\end{equation}

The fixed polynomial
\[
 G_N(T)=H_N(T)+T=T^N+T^3-2T^2+T
\]
satisfies the hypotheses of Proposition~\ref{prop:quadratic}.
Its quadratic-gauge map recovers \(P_{N,u}\) at
\begin{equation}
 q_u^{\rm quad}=(1,0,-2u).
 \label{eq:quadratic-target-N}
\end{equation}

\begin{lemma}[Boundary cleanliness]\label{lem:boundary-clean}
The seed \(H_N\) is boundary-clean for every \(N\geq4\).
\end{lemma}

\begin{proof}
One has
\[
 H_N(T)=T^2(T^{N-2}+T-2).
\]
The last factor is nonzero at zero and vanishes at one.  If it shared a
root with its derivative, then
\[
 T^{N-3}=-\frac1{N-2}
\]
and substitution into \(T^{N-2}+T-2=0\) would give
\[
 T=\frac{2(N-2)}{N-3}>0,
\]
a contradiction.
\end{proof}

\begin{lemma}[Generic connectedness]\label{lem:generic-irreducible}
The polynomial \(P_{N,U}\) is irreducible over \(\Q(U)\).
\end{lemma}

\begin{proof}
It is primitive in \(\Q[T,U]\) and has degree one in \(U\).  In any
factorization, one factor has \(U\)-degree zero.  Comparing the coefficient
of \(U\) makes that factor a unit in \(\Q[T]\).  Gauss's lemma finishes the
proof.
\end{proof}

Hilbert irreducibility applied to Lemma~\ref{lem:generic-irreducible} gives
a Hilbert subset of rational \(u\) with \(P_{N,u}\) irreducible
~\cite[Chapter~3]{Serre}.  Propositions~\ref{prop:weighted} and
\ref{prop:quadratic} give the two complete fibers, while
Lemma~\ref{lem:boundary-clean} permits the stable unit-rank separation in
Section~\ref{sec:stable}.  This proves Theorem~\ref{thm:rational-pairs}.

\begin{proposition}[A small rational quartic]\label{prop:small-Q}
Let
\[
 P(T)=2T^4-T^3-T^2+T+1,\qquad A=\Q[T]/(P).
\]
Then \(A\) is a quartic field and \(|\cR_\Q(A)|\geq2\).
\end{proposition}

\begin{proof}
The tangent data are
\[
 P(1)-P(0)=P'(0)=1,\quad P'(0)-P'(1)=-3,
\]
\[
 P''(1)=16\neq6=2(P'(1)-P'(0)).
\]
The weighted seed
\[
 P(T)-1-T=T^2(T-1)(2T+1)
\]
is boundary-clean, and the weighted target is
\((-1/3,-1,1)\).  The quadratic-gauge seed \(P-1\) has
\((g_1,g_3,g_4)=(1,-1,2)\), and target \((1,0,-2)\).

Modulo \(3\), multiplication by the inverse leading coefficient gives
\[
 \overline P=T^4+T^3+T^2-T-1.
\]
It has no root in \(\mathbb F_3\).  The only monic irreducible quadratic
over \(\mathbb F_3\) is \(T^2+1\), and division leaves remainder \(T-1\).
Thus \(P\) is irreducible.  Its discriminant is \(1556\), so the common
fiber is connected and etale.
\end{proof}

\section{Three fixed maps with infinitely many common fibers}

Let \(K=\Q(\eta)\), \(\eta^2=-2\), and consider \(\Psi_R\) from
\eqref{eq:triple-pencil}.  It is already the inverse polynomial of a fixed
quartic cancellation map.

For the type-\((2,1)\) cancellation construction, put
\[
 A_0=1+xy^2,\qquad
 B_0=A_0^2z+y^3h(A_0),\qquad
 \Pi=A_0B_0,\qquad Q=y+xB_0,
\]
and
\begin{equation}
 \mathcal R=\int_0^{x/A_0}\left(1-t(Q-\Pi t)^2\right)\,dt.
 \label{eq:cancellation-map}
\end{equation}
The apparent denominator in \(\mathcal R\) cancels precisely when the
finite jet \(h\) satisfies the cancellation equation.  Take
\[
 \theta=2+\eta,\qquad
 h(A_0)=\theta+(4\theta-6)A_0.
\]
Since \(\theta^2-4\theta+6=0\), the map
\[
 F^{\rm can}=(\Pi,Q,\mathcal R)
\]
is polynomial and has determinant \(-1\).  Its inverse equation at a target
\((\Pi,Q,R)\) is
\begin{equation}
 T-\frac{Q^2}{2}T^2+\frac{2\Pi Q}{3}T^3
 -\frac{\Pi^2}{4}T^4-R.
 \label{eq:cancellation-inverse}
\end{equation}
At \((\Pi,Q,R)=(4+\eta,3,R)\), equation
\eqref{eq:cancellation-inverse} is exactly \(\Psi_R\).

The same pencil is tangent-admissible:
\[
 \Psi_R(1)-\Psi_R(0)=\Psi_R'(0)=1,
\]
\[
 \Psi_R'(0)-\Psi_R'(1)=-1+2\eta,
\]
\[
 \Psi_R''(1)-2(\Psi_R'(1)-\Psi_R'(0))=-5-8\eta.
\]
Its weighted seed is fixed and boundary-clean:
\begin{equation}
 H(T)=T^2(T-1)
 \left(\frac92-\left(\frac72+2\eta\right)T\right).
 \label{eq:triple-weighted-seed}
\end{equation}
The corresponding weighted target is
\begin{equation}
 q_R^{\rm wt}=
 \left(\frac{-R}{-1+2\eta},-1,1\right).
 \label{eq:triple-weighted-target}
\end{equation}

The fixed quadratic-gauge seed is
\[
 G(T)=\Psi_R(T)+R,
\]
whose coefficients satisfy \(g_1g_3g_4\neq0\).  Its target is
\begin{equation}
 q_R^{\rm quad}=(1,0,2R).
 \label{eq:triple-quadratic-target}
\end{equation}
The reconstruction derivatives in all three constructions are nonzero
exactly when \(\Psi_R\) is squarefree.  Thus every such specialization gives
the same complete finite etale \(K\)-algebra.

As a polynomial over \(K(R)\), \(\Psi_R\) is irreducible by the same
degree-one-in-\(R\) argument as Lemma~\ref{lem:generic-irreducible}.
Hilbert irreducibility gives the connected specializations.  The stable
separation in Section~\ref{sec:stable} now proves
Theorem~\ref{thm:quadratic-triples}.

\begin{proposition}[An explicit connected triple fiber]
\label{prop:mod17}
At \(R=-1\), the algebra \(K[T]/(\Psi_{-1})\) is a quartic field.
\end{proposition}

\begin{proof}
Reduce \(2\Psi_{-1}\) at the degree-one prime
\((17,\eta-10)\).  Since \(10^2\equiv-2\pmod {17}\), this is a valid prime
of \(K\).  After monic normalization the reduction is
\[
 f(T)=T^4+14T^3+2T^2+9T+9\in\mathbb F_{17}[T].
\]
Exact modular powering gives
\[
 T^{17^4}-T\equiv0\pmod f
\]
and
\[
 T^{17^2}-T\equiv6T^3-T^2-6T+3\pmod f,\qquad
 \gcd(f,T^{17^2}-T)=1.
\]
Rabin's criterion proves that \(f\) is irreducible.  Hence
\(\Psi_{-1}\) is irreducible over \(K\); characteristic zero gives
separability.
\end{proof}

\section{Stable inequivalence}\label{sec:stable}

We summarize the intrinsic boundary calculation used above.  For a
dominant quasi-finite map \(F:U\to Y\), let
\[
 \overline X_F=\Norm_Y K(U)
\]
and retain the open immersion \(U\hookrightarrow\overline X_F\) and its
reduced boundary.  This canonical finite-normalization package is functorial
under polynomial source and target automorphisms.  Identity stabilization
replaces it by its product with affine space.  Consequently the unit group
of an intrinsic normalized stratum, relative Fitting ideals, and
scheme-theoretic boundary-contact nilpotency are stable left-right
invariants.  Full proofs and exact boundary ledgers appear in
~\cite{StableRepository}.

For the three families used here, the boundary package intrinsically orders
two target images:
\[
 Z_\Delta=\text{the vertex receiving a ramified boundary prime},\qquad
 Z_0=\text{the other vertex}.
\]
After deleting \(Z_0\), the normalized ramified target strata and their
unit ranks are
\begin{center}
\begin{tabular}{lcc}
\toprule
construction&
\(\Norm(Z_\Delta\setminus Z_0)\)&
\(\rank(\mathcal O^\times/K^\times)\)\\
\midrule
weighted boundary-clean&\(\A^1\times\Gm\)&\(1\)\\
cancellation type \((2,1)\)&\(\Gm^2\)&\(2\)\\
quadratic gauge&\(\Gm^2\)&\(2\)\\
\bottomrule
\end{tabular}
\end{center}
Affine-space factors introduce no nonconstant units.  Unit rank therefore
separates every weighted map used here from either reciprocal map.

For cancellation, write torus coordinates \((Y,v)\) with
\(Q=(v+1)Y\).  Up to a Laurent monomial unit, the relative Fitting generator
is
\[
 J_{\rm can}=2v-1.
\]
Its Laurent support has affine rank one.  For a quartic quadratic gauge,
after the standard shear and scaling, the corresponding generator contains
the exponent points
\[
 (0,0),\qquad(1,2),\qquad(4,3),
\]
which have affine rank two.  Torus automorphisms act on exponent lattices
through affine \(\mathrm{GL}_2(\mathbb Z)\) transformations, preserving this
rank.  Thus cancellation and quadratic gauge cannot become equivalent after
stabilization.

There is also an independent contact check.  The nilpotency indices of the
scheme-theoretic intersection \(Z_\Delta\cap Z_0\) are
\begin{equation}
 \mu_{\rm wt}=1,\qquad
 \mu_{\rm can}=6,\qquad
 \mu_{\rm quad}=2.
 \label{eq:contact-indices}
\end{equation}
Polynomial extension preserves the least nilpotency exponent.  Equation
\eqref{eq:contact-indices} independently separates all three constructions.

\section{Normalization, interpretation, and scope}

If a displayed map has determinant \(d\), translate its distinguished target
to the origin and multiply one target coordinate by \(d^{-1}\).  This target
automorphism changes the determinant to one, preserves the complete fiber,
and leaves the left-right class unchanged.  All maps in
Theorems~\ref{thm:rational-pairs} and \ref{thm:quadratic-triples} may
therefore be presented with determinant one and common fiber over the
origin.

The point is not that two polynomials happen to have the same root set.
The quotient-ring reconstruction in Propositions~\ref{prop:weighted} and
\ref{prop:quadratic}, and its cancellation analogue, identifies the entire
scheme fiber with the same finite etale algebra.  The surrounding maps are
then separated by invariants of their canonical behavior at infinity.

The public literature currently treats arithmetic images, individual
fibers, and generic degrees of Keller maps from several directions; see
Gallagher's exact degree atlas~\cite{Gallagher}, the graded arithmetic
analysis of Shaska~\cite{Shaska}, the ideal-theoretic fiber result of
Miranda-Neto~\cite{Miranda}, and the Hilbert-irreducibility viewpoint of
Lipton and Markakis~\cite{LiptonMarkakis}.  The results here instead fix the
finite etale algebra and ask how many stable Keller geometries contain it.
No priority claim is made beyond the dated sources cited here.

\paragraph{Verification.}
The repository checker
\[
 \texttt{scripts/verify\_common\_arithmetic\_fibers.py}
\]
replays the tangent identities, boundary cleanliness, common inverse
polynomials, the small rational quartic, the quadratic-field cancellation
specialization, and the mod-\(17\) Rabin certificate.  Independent scripts
replay the three family-specific stable boundary calculations.

\paragraph{AI assistance.}
Generative AI tools were used for symbolic experimentation, exact checker
development, adversarial proof review, and manuscript preparation.  The
author reviewed the mathematical claims and assumes full responsibility for
the contents.

\begin{thebibliography}{9}
\small

\bibitem{Gallagher}
A.~Gallagher,
\emph{Generic fiber degrees 3 through 100},
online exact certificate atlas, 2026,
\href{https://jacobianfun.org/counterexamples}{jacobianfun.org}.

\bibitem{LiptonMarkakis}
R.~J. Lipton and E.~Markakis,
\emph{Some remarks on the Jacobian conjecture and connections with
Hilbert's irreducibility theorem},
\href{https://arxiv.org/abs/math/0507525}{arXiv:math/0507525}, 2005.

\bibitem{Miranda}
C.~B. Miranda-Neto,
\emph{An ideal-theoretic approach to Keller maps},
Proc. Edinburgh Math. Soc. \textbf{62} (2019), 1033--1044,
\href{https://doi.org/10.1017/S0013091519000099}
{doi:10.1017/S0013091519000099}.

\bibitem{Serre}
J.-P. Serre,
\emph{Topics in Galois Theory},
2nd ed., Research Notes in Mathematics 1,
A K Peters, 2008.

\bibitem{Shaska}
T.~Shaska,
\emph{Graded Keller maps and the Jacobian conjecture},
\href{https://arxiv.org/abs/2607.20210}{arXiv:2607.20210}, 2026.

\bibitem{StableRepository}
R.~van Rijn,
\emph{From one collision to marked-root Keller maps},
research repository, version of 24 July 2026,
\href{https://github.com/royvanrijn/jacobian-research}{GitHub repository}.

\end{thebibliography}

\end{document}
