\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,microtype}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newcommand{\A}{\mathbb A}
\newcommand{\Pone}{\mathbb P^1}

\title{Discriminants of the Pencils $H(W)-sW+t$}
\author{Repository working draft}
\date{\today}

\begin{document}
\maketitle

\begin{abstract}
Let $k$ be an algebraically closed field of characteristic zero and let
$H\in k[W]$ have degree $n\geq3$.  The reduced discriminant of the pencil
$H(W)-sW+t$ is parameterized by
\[
 r\longmapsto \bigl(H'(r),rH'(r)-H(r)\bigr).
\]
We prove that for $H$ in a nonempty Zariski-open set its projective closure
is a rational plane curve of degree $n$, smooth at its unique point at
infinity, with exactly $n-2$ ordinary cusps and
$\frac12(n-2)(n-3)$ ordinary nodes, and with no other singularities.  The
same open set meets a natural normalized endpoint slice in every degree.
The proof uses contact factorizations and dimension estimates; the genus
formula is used only after all possible singularity types have been
excluded.
Over the reals we also prove a complementary chamber theorem: if one member
of the pencil has $n$ simple real roots, then its chambers realize exactly
$n,n-2,\ldots,n\bmod2$ real roots.  A single generic vertical line meets
these chambers successively across ordinary folds, and rational input data
admit rational witnesses in every chamber.
\end{abstract}

\section{Introduction}

Pencils of the form
\[
 E_{s,t}(W)=H(W)-sW+t                                      \tag{1.1}
\]
occur when a polynomial inverse problem is reduced to choosing a marked
root.  In the motivating Jacobian-map construction, the target coordinates
specialize to $s=BC$ and $t=cAC^2$.  None of that construction is needed
below: our subject is the intrinsic discriminant geometry of (1.1).

The common organizing map is the
\href{../../verified/TANGENT_MAP_CORE.md}{tangent-map core}.  In its normal
coordinates it is
\[
 (W,s)\longmapsto\bigl(s,Ws-H(W)\bigr),
\]
whose inverse equation is (1.1).  Its critical restriction gives at once
\[
 s=H'(r),\qquad t=rH'(r)-H(r).                             \tag{1.2}
\]
Thus the discriminant records tangent lines to the affine graph $Y=H(W)$.
Its cusps correspond to flexes and its self-intersections to bitangent
lines.  Rather than invoke a general theorem on generic dual curves, we use
the special form (1.2) to control all bad contacts, including their behavior
at infinity.

Throughout, ``the discriminant curve'' means the reduced projective closure
of the repeated-root locus.  Adding an affine-linear polynomial to $H$
only applies an affine transformation to $(s,t)$, so we work in
\[
 \mathcal B_n=\operatorname{Spec}k[h_2,\ldots,h_n],\qquad
 H(W)=\sum_{j=2}^n h_jW^j,                                 \tag{1.3}
\]
and write $\mathcal B_n^\circ=D(h_n)$ for the exact-degree locus.

\section{Statement of the theorem}

\begin{theorem}\label{thm:main}
For every $n\geq3$ there is a nonempty Zariski-open subset
$\mathcal U_n\subset\mathcal B_n^\circ$ such that for every
$H\in\mathcal U_n$ the discriminant curve of $H(W)-sW+t$ has:
\begin{enumerate}
\item degree $n$ and normalization $\Pone$;
\item one point at infinity, which is smooth;
\item exactly $n-2$ ordinary cusps;
\item exactly $\frac12(n-2)(n-3)$ ordinary nodes;
\item no other singularities.
\end{enumerate}
Moreover, $\mathcal U_n$ meets the normalized endpoint slice
\[
 \mathcal A_n=\{H:\deg H=n,\ H(0)=H'(0)=H(1)=0,
                   \ H'(1)=-1,\ H''(1)\ne-2\}.             \tag{2.1}
\]
Consequently the same conclusions hold on a nonempty open subset of
$\mathcal A_n$.
\end{theorem}

The characteristic-zero assumption is used for derivatives, separability,
and the local classification of the parameterized branches.  The theorem is
not asserted in small positive characteristic.

\section{Normalization, degree, and infinity}

Homogenizing (1.2) gives a morphism
\[
 \bar\nu_H:\Pone\longrightarrow\mathbb P^2,
 \qquad [R:U]\longmapsto[S_H:T_H:Z],                       \tag{3.1}
\]
where
\begin{align*}
 S_H&=\sum_{j=2}^n jh_jR^{j-1}U^{n-j+1},\\
 T_H&=\sum_{j=2}^n (j-1)h_jR^jU^{n-j},\\
 Z&=U^n.
\end{align*}
These degree-$n$ forms have no common zero on $D(h_n)$.  The only parameter
above the line at infinity is $[1:0]$, and its image is $[0:1:0]$.

\begin{proposition}\label{prop:normalization}
For every $H\in\mathcal B_n^\circ$, the map $\bar\nu_H$ is finite and
birational onto an integral plane curve of degree $n$.  It is a formal
closed immersion at $[1:0]$; in particular its image is smooth there and no
off-diagonal equal-image pair can specialize to that point while $h_n$
remains nonzero.
\end{proposition}

\begin{proof}
On the affine chart,
\[
 \frac{ds}{dr}=H''(r),\qquad \frac{dt}{dr}=rH''(r).
\]
At the generic point of the image, $r=dt/ds$; hence the map is birational.
A nonconstant map from the proper curve $\Pone$ to its image is finite.
The base-point-free forms in (3.1) pull a general line back to a divisor of
degree $n$.  Birationality therefore makes the image degree $n$.

For the assertion at infinity, put $q=U/R$ and use the chart $T_H\ne0$.
Uniformly over $k[h_2,\ldots,h_n,h_n^{-1}]$,
\[
 x:=\frac{S_H}{T_H}=\frac n{n-1}q+O(q^2),\qquad
 y:=\frac Z{T_H}=\frac{q^n}{(n-1)h_n}+O(q^{n+1}).           \tag{3.2}
\]
The first series has unit linear coefficient, so the formal inverse theorem
recovers $q$ uniquely from $x$.  The completed image is therefore the graph
$y=f(x)$, proving formal closed immersion and smoothness.

For completeness, consider the relative fiber product of two copies of
$\Pone$ over the image while the coefficients vary in $D(h_n)$.  If an
off-diagonal pair approached the infinity section, the uniqueness of the
preimage of $Z=0$ forces both parameters to approach $q=0$.  In the completed
neighborhood, equality of the two $x$-coordinates and formal invertibility
in (3.2) force $q_1=q_2$.  Thus the off-diagonal locus has closure disjoint
from the infinity section.
\end{proof}

\section{Contact strata and exclusion of bad singularities}

Let $\ell_r$ denote the tangent line to $Y=H(W)$ at $r$, and set
\[
 c(r)=\operatorname{ord}_{W=r}\bigl(H(W)-\ell_r(W)\bigr)\geq2.
\]
Differentiation of (1.2) gives
\[
 \nu_H'(r)=H''(r)(1,r).                                    \tag{4.1}
\]
Thus $c(r)=2$ gives a smooth immersed branch.  If $c(r)=3$, then
$H'''(r)\ne0$ and
\[
 \det\bigl(\nu_H''(r),\nu_H'''(r)\bigr)=2H'''(r)^2\ne0,   \tag{4.2}
\]
so the image is an ordinary cusp.  Contact $c(r)\geq4$ is the first bad
one-branch event.

If two distinct parameters $r,u$ have the same image, the graph has a common
tangent line at both points.  When both branches are unramified their tangent
directions are $(1,r)$ and $(1,u)$, with determinant $u-r\ne0$.  The
intersection is automatically an ordinary transverse node.  The same
determinant is the Jacobian of the equal-image equations with respect to the
two parameters, so the tangent--chord coincidence is transverse as an
incidence as well.  In particular a tacnode cannot occur from two distinct
unramified normalization points.

It remains to show that all other events are nongeneric.  For a contact
pattern $\mu=(\mu_1,\ldots,\mu_k)$, write
\[
 M_\mu(W)=\prod_{i=1}^k(W-r_i)^{\mu_i},\qquad
 m=\sum_i\mu_i.
\]
A line with these contacts gives the factorization
\[
 H(W)-\ell(W)=M_\mu(W)Q(W),\qquad \deg Q\le n-m.           \tag{4.3}
\]
After discarding the affine-linear coefficients, (4.3) defines a polynomial
map to $\mathcal B_n$ from a space of dimension
\[
 k+(n-m+1).                                                 \tag{4.4}
\]
For each of the three patterns
\[
 (4),\qquad(3,2),\qquad(2,2,2),                            \tag{4.5}
\]
this dimension equals $n-2$ whenever the pattern is possible.  The Zariski
closure of the image of a finite-type morphism cannot have larger dimension
than its source.  Hence the closed image $Z_\mu\subset\mathcal B_n$ of every
pattern in (4.5) has codimension at least one.  This argument already takes
the closure: collisions among marked roots, roots of $Q$ meeting a marked
root, and degree drops in $Q$ cannot create a larger boundary component.

\begin{lemma}\label{lem:exhaustion}
Outside $Z_{(4)}\cup Z_{(3,2)}\cup Z_{(2,2,2)}$, every affine singularity of
the discriminant curve is an ordinary node or ordinary cusp.
\end{lemma}

\begin{proof}
A nonordinary single branch has contact at least four and is in $Z_{(4)}$.
If a fiber of the normalization contains two points and one branch is
ramified, the contact pattern dominates $(3,2)$.  If it contains at least
three points, any three give $(2,2,2)$.  These statements include higher
contact, cusp--cusp collisions, tritangents, and two nominal node pairs with
the same image.  The only remaining case has exactly two unramified
preimages, and (4.1) makes their branches transverse.
\end{proof}

The proposition at infinity is essential here: it prevents a missing
projective bad stratum.  A ramification point cannot move to parameter
infinity on $D(h_n)$ because the homogenization of $H''$ there has value
$n(n-1)h_n$.  Nor can a finite branch share the image $[0:1:0]$.  Finally,
the relative fiber-product statement in Proposition~\ref{prop:normalization}
rules out two finite equal-image parameters escaping together to infinity.
Thus (4.5) exhausts the projective, not merely affine, failure modes.

The complement
\[
 \mathcal U_n=\mathcal B_n^\circ\setminus
 (Z_{(4)}\cup Z_{(3,2)}\cup Z_{(2,2,2)})                  \tag{4.6}
\]
is nonempty and open.  On it, $H''$ is squarefree, so its $n-2$ roots give
$n-2$ distinct ordinary cusps.  No cusp shares its image with another
branch, and every other singularity is an ordinary node.

\section{The endpoint slice is not trapped in the bad locus}

We now prove the last genericity assertion in Theorem~\ref{thm:main}.  It is
not enough to know that (4.6) is nonempty in the full coefficient space.

For $G\in\mathcal B_n$ and $\alpha\ne\beta$, define the divided
tangent--chord equation
\[
 \Phi(G,\alpha,\beta)=
 \frac{G(\beta)-G(\alpha)-G'(\alpha)(\beta-\alpha)}
      {(\beta-\alpha)^2}=0.                                \tag{5.1}
\]
It is polynomial, and the coefficient of $h_2$ in $\Phi$ is one.  Therefore
the incidence $\mathcal T_n=V(\Phi)$ over $\alpha\ne\beta$ is an
irreducible affine bundle.  Its projection to $\mathcal B_n^\circ$ is
surjective: for any exact-degree $G$, choose $\alpha$ with
$G''(\alpha)\ne0$.  As a polynomial in $\beta$, (5.1) has degree $n-2$ and
does not vanish at $\beta=\alpha$, so it has a root distinct from $\alpha$.

Put $d=\beta-\alpha$ and
\[
 H(W)=G(\alpha+dW)-G(\alpha)-G'(\alpha)dW.                 \tag{5.2}
\]
Equation (5.1) gives $H(0)=H'(0)=H(1)=0$.  Source
reparameterization and an invertible affine transformation of $(s,t)$ carry
the discriminant of $G$ to that of $H$, preserving all projective
singularity types, including the point at infinity.  Hence the inverse image
of $\mathcal U_n$ in $\mathcal T_n$ is a nonempty open.

The additional conditions
\[
 \deg H=n,\qquad H'(1)\ne0,\qquad H''(1)-2H'(1)\ne0         \tag{5.3}
\]
also define a nonempty open in $\mathcal T_n$.  A uniform witness is
\[
 H_n(W)=W^2+W^3+\cdots+W^{n-1}-(n-2)W^n,\qquad
 (\alpha,\beta)=(0,1),                                    \tag{5.4}
\]
for which $H_n(1)=0$, $H_n'(1)\ne0$, and
$H_n''(1)/(-H_n'(1))=-4n/3\ne-2$.  Since $\mathcal T_n$ is irreducible, the
two nonempty opens intersect.  Scaling (5.2) by $-1/H'(1)$ reaches
$H'(1)=-1$ without changing any singularity type or the last inequality in
(5.3).  This proves that $\mathcal A_n\cap\mathcal U_n$ is nonempty.

\section{Counting the nodes}

Only now do we use the genus formula.  For $H\in\mathcal U_n$, the curve is
integral of degree $n$ with normalization $\Pone$.  Sections 3 and 4 have
already excluded every singularity except ordinary nodes and the $n-2$
ordinary cusps.  Both types have delta invariant one.  If $N$ denotes the
number of nodes, then
\[
 \frac{(n-1)(n-2)}2
 =p_a(C)=g(\Pone)+\sum_{x\in\operatorname{Sing}C}\delta_x
 =0+(n-2)+N.
\]
Thus
\[
 \boxed{N=\frac{(n-2)(n-3)}2}.
\]
The genus formula counts the nodes left after the geometric exclusion; it is
not used to infer that unclassified singularities must be nodes.

\section{The real chamber spectrum}

We finish with a real statement which does not require the genericity open
$\mathcal U_n$.  Let
\[
 U_H(\mathbb R)=
 \{(s,t)\in\mathbb R^2:\operatorname{Disc}_W(H(W)-sW+t)\ne0\}.
\]
The number of real roots is locally constant there and hence constant on
each chamber.

\begin{theorem}\label{thm:real-spectrum}
Let $H\in\mathbb R[W]$ have degree $n\ge2$.  If some member of
$H(W)-sW+t$ has $n$ distinct real roots, then the real-root counts occurring
on the chambers of $U_H(\mathbb R)$ are exactly
\[
 n,n-2,n-4,\ldots,n\bmod2.                               \tag{7.1}
\]
Arbitrarily close to the slope of the split member there is one vertical
line in the $(s,t)$-plane which encounters all these counts successively,
with each consecutive pair separated by one smooth fold.  If $H$ and the
split member are defined over $\mathbb Q$, that slope and a point in every
chamber in this chain may be chosen rational.
\end{theorem}

\begin{proof}
Complex conjugation pairs the nonreal roots of every squarefree real member,
so a real-root count $r$ necessarily satisfies $0\le r\le n$ and
$r\equiv n\pmod2$.

After translating $t$, suppose the split member occurs at $(s_0,0)$ and put
$P_s(W)=H(W)-sW$.  Having $n$ simple real roots is open in $s$.  In a small
interval about $s_0$, avoid the slopes $H'(u)$ with $H''(u)=0$ and the slopes
at which two distinct critical points have equal critical values.  The first
set is finite.  The second is finite as well: equal critical values are
distinct normalization points with the same image under
\[
 u\longmapsto\bigl(H'(u),uH'(u)-H(u)\bigr),              \tag{7.2}
\]
and the finite birational normalization map (7.2) is one-to-one away from
finitely many points.  Denote a remaining slope by $s_1$.  Rolle's theorem
shows that $P_{s_1}$ has exactly $n-1$ real critical points; they are
nondegenerate and their critical values are pairwise distinct.

Let $a$ be the leading coefficient of $H$ and $d=\operatorname{sign}(a)$.
For $q\ge0$ consider
\[
 P_{s_1}(W)+dq.                                          \tag{7.3}
\]
After multiplying by $-1$ if necessary, the leading coefficient is positive
and (7.3) raises the polynomial through its negative local minima.  Because
$P_{s_1}$ has $n$ simple real zeros, every local minimum is negative and
every local maximum is positive.  The relevant critical levels are positive
and distinct.  Crossing each one annihilates precisely two real roots.  For
large $q$ there are no real roots in even degree and one in odd degree.
This gives every count in (7.1) in order, and the nondegenerate critical
points give smooth folds.

For rational data, the split-root condition holds on a real interval of
slopes.  The finitely many excluded slopes leave a rational $s_1$.  The
successive critical levels are distinct real algebraic numbers, so every
open interval between them contains a rational $q$.  Equation (7.3) then
gives the asserted rational witnesses.
\end{proof}

\begin{remark}[Exact chamber graph]
For a fixed $H\in\mathbb Q[W]$, the entire adjacency graph is algorithmic.
Form $D_H(s,t)=\operatorname{Disc}_W(H(W)-sW+t)$ and perform cylindrical
algebraic decomposition with the leading coefficients and
$\operatorname{Disc}_t(D_H)$ among the projection polynomials.  Between
successive event slopes, exact real-root isolation orders the critical
$t$-values and Sturm counting labels every band.  Gluing across the event
fibers gives the dual chamber graph.  Theorem~\ref{thm:real-spectrum}
extracts from it the universal path (7.1); additional same-count chambers
and node-induced adjacencies remain seed-dependent.
\end{remark}

\section*{References}

The proof uses only standard facts about finite maps of proper curves,
dimension of images, formal inversion, and the normalization-genus formula.
For proper images see the Stacks Project, Tag
\href{https://stacks.math.columbia.edu/tag/01W6}{01W6}.  For broader context
on inflections and bitangents as singularities of dual curves, see
A.~Yuran, \emph{Pl\"ucker formulas for plane algebraic curves with a given
Newton polygon}, arXiv:\href{https://arxiv.org/abs/2204.04626}{2204.04626}.

\end{document}
