\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,microtype}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\newcommand{\A}{\mathbb A}
\renewcommand{\P}{\mathbb P}
\newcommand{\C}{\mathbb C}

\title{Decorated Discriminant Normalization and Stable Moduli}
\author{Repository working draft}
\date{\today}

\def\DecoratedNormalizationPaper{1}

\begin{document}
\maketitle

\begin{abstract}
For the weighted marked-root Keller maps attached to a normalized seed $H$,
we study the normalization of the discriminant curve of the pencil
$H(W)-sW+t$.  Its zeroth relative Fitting ideal is $(H''(r))$ with full
multiplicity.  The marked Hessian divisor
$(\P^1;\operatorname{div}(H''),0,\infty)$ is the minimal stable invariant
needed for generic moduli: its quotient map is generically etale of exact
degree $N-2$ onto an image of dimension $N-3$.  Adding the unique unramified
affine root sheet recovers the normalized seed.  The saturated off-diagonal
node-pairing scheme and conductor morphism are retained separately as the
stronger scheme-theoretic D2 decoration; they are not used in the generic
dimension, unramifiedness, or rerooting-degree proof.
\end{abstract}

\section{Setup and scope}

Let $k$ be a characteristic-zero field and let $H\in k[W]$ be a normalized
admissible degree-$N$ seed:
\[
 H(0)=H'(0)=H(1)=0,\qquad H'(1)=-1.
\]
The construction-independent
\href{../../verified/TANGENT_MAP_CORE.md}{tangent-map core} is the common
opening.  In normal coordinates
\[
 (W,s)\longmapsto\bigl(s,Ws-H(W)\bigr),
\]
it gives simultaneously the inverse pencil, its discriminant normalization,
the Hessian Fitting divisor, and the critical-value incidence used below.
The weighted map is its suspension.

The weighted construction associates to $H$ a polynomial Keller map $F_H$
and the marked inverse equation
\[
 E_H(W;s,t)=H(W)-sW+t=0.
\]
Its finite incidence normalization is equipped with the maximal open on which
the rational source reconstruction is regular.  The construction-independent
\href{../../verified/STABLE_NORMALIZATION_FUNCTORIALITY.md}
{stable-normalization functoriality theorem} identifies the
intrinsic discriminant vertex, its finite normalization morphism, and its
boundary marks under polynomial left--right equivalence and after adjoining
identity variables.  Those facts are the functorial input used here.

The marked Hessian-divisor quotient deliberately forgets the affine root
sheet, so its generic degree is the rerooting degree $N-2$.  Retaining the
unique unramified sheet meeting the regular-reconstruction open removes
exactly this ambiguity and gives generic faithfulness.  Node pairing and the
conductor refine this quotient to the full D2 decoration, but play no role in
D1 or F2.  The companion Hurwitz--LL theorem
retains the selected root on the marked admissible-cover stack and identifies
the normalized Stein factors, completed charts, and conductor squares at
arbitrary collisions.  The finite-birational argument on the normal marked
coarse compactification also descends that root to a unique coarse affine
mark over every DVR limit.  At a multiplicity-$\mu$ collision its fiber is
$k[T]/(T^\mu)$, nonreduced but supported at one geometric point.

\input{../marked-root-multiplicity/decorated-normalization.tex}

\end{document}
