\section{The explicit Keller map}\label{sec:explicit}

Throughout Sections~\ref{sec:explicit}--\ref{sec:cubic-image}, let $k$ be a
field of characteristic zero.  Write $\A^m_k$ for affine $m$-space over $k$.
Set
\[
 u=1+xy
\]
and define $F=(F_1,F_2,F_3):\A^3_k\to\A^3_k$ by
\begin{align}
F_1&=u^3z+y^2u(4+3xy),\nonumber\\
F_2&=y+3xu^2z+3xy^2(4+3xy),\label{eq:F}\\
F_3&=2x-3x^2y-x^3z.\nonumber
\end{align}

\begin{theorem}[constant Jacobian and collision]\label{thm:counterexample}
The morphism $F$ is etale and
\[
 \det DF=-2.
\]
Moreover the three distinct $k$-points
\[
p_0=(0,0,-1/4),\quad p_+=(1,-3/2,13/2),\quad
p_-=(-1,3/2,13/2)
\]
have the common image $(-1/4,0,0)$.  Hence $F$ is not injective and is not a
polynomial automorphism.  For every $d\ge3$, the product
$F\times\operatorname{id}_{\A^{d-3}}$ has the same properties.
\end{theorem}

\begin{proof}
On the dense open subset $D(x)\subset\A^3_k$, put
\[
 t=y+x^{-1},\qquad r=2x^{-1},\qquad c=F_3(x,y,z),
\]
and write $(a,b,c)=F(x,y,z)$.  Direct expansion in
$k[x,x^{-1},y,z]$ gives
\begin{equation}\label{eq:trc-target}
 a=t^2+\frac{rt}{2}-ct^3,qquad b=r+4t-3ct^2.
\end{equation}
Consequently
\[
 \det\frac{\partial(a,b,c)}{\partial(t,r,c)}=\frac r2,
 \qquad
 \det\frac{\partial(t,r,c)}{\partial(x,y,z)}=-2x.
\]
Since $r=2/x$, the chain rule gives $\det DF=-2$ on $D(x)$.
The determinant is a polynomial, and $D(x)$ is dense because $k$ is an
infinite field, so the identity holds on all of $\A^3_k$.  Since $-2$ is a
unit, the Jacobian criterion makes $F$ etale.

Substitution into \eqref{eq:F} gives
\[
 F(p_0)=F(p_+)=F(p_-)=(-1/4,0,0).
\]
The three first coordinates $0,1,-1$ show that the source points are
distinct.  The final assertion follows because the determinant of a product
map is the product of the determinants and the displayed collision persists
after fixing the additional coordinates.
\end{proof}

The proof above is entirely algebraic and applies over every
characteristic-zero field.  Independent exact implementations, including a
separately authored Lean formalization, verify the determinant and collision;
they are verification artifacts rather than inputs to the argument.
