\section{Fibers, image, and nonproperness}\label{sec:cubic-image}

For this section, assume first that $k$ is algebraically closed of
characteristic zero.  Put
\begin{equation}\label{eq:Qdisc}
 \Delta(a,b,c)=27a^2c^2-18abc+16a+b^3c-b^2
\end{equation}
and
\begin{equation}\label{eq:Gamma}
 \Gamma=V(3bc-4,\,12a-b^2)\subset\A^3_k.
\end{equation}
The discriminant of $P(T)=cT^3-2T^2+bT-2a$ is $-4\Delta$.

\begin{proposition}[geometric fibers and image]\label{prop:fibers-image}
For every $q=(a,b,c)\in\A^3_k$,
\[
 \#F^{-1}(q)=
 \begin{cases}
 3,&\Delta(q)\ne0,\\
 1,&\Delta(q)=0\text{ and }q\notin\Gamma,\\
 0,&q\in\Gamma.
 \end{cases}
\]
Consequently $F(\A^3_k)=\A^3_k\setminus\Gamma$.
\end{proposition}

\begin{proof}
By Theorem~\ref{thm:marked-root}, the fiber is the set of simple projective
roots of the nonzero binary cubic \eqref{eq:binary-cubic}.  A binary cubic
over an algebraically closed field has multiplicity type $(1,1,1)$,
$(2,1)$, or $(3)$, yielding respectively three, one, or zero simple roots.
The first type is equivalent to $\Delta\ne0$.  A comparison of coefficients
with a cube shows that type $(3)$ occurs exactly when
$3bc=4$ and $12a=b^2$; these equations force $c\ne0$.  This is precisely
$\Gamma$, and all remaining discriminant points have type $(2,1)$.
\end{proof}

We now specialize to $k=\mathbb C$.  For a polynomial map
$f:\mathbb C^m\to\mathbb C^m$, define its nonproperness set $S_f$ to be the
set of targets $q$ for which there are $p_j\in\mathbb C^m$ with
$\lVert p_j\rVert\to\infty$ and $f(p_j)\to q$.

\begin{proposition}[exact nonproperness set]\label{prop:nonproperness}
For the map $F$ in \eqref{eq:F},
\[
 S_F=V(\Delta)\subset\mathbb C^3.
\]
\end{proposition}

\begin{proof}
Let $q_0=(a_0,b_0,c_0)\in V(\Delta)$.  Its cubic has a repeated finite root
$t_0$: a repeated root cannot be the projective root $[1:0]$, since on the
$s=V/U$ chart one has $R'(0)=-2$.  Keep $b=b_0$, $c=c_0$, let $t\to t_0$,
and set
\[
 a(t)=\frac{ct^3-2t^2+bt}{2}.
\]
Then $P_t(t)=0$, the targets $(a(t),b,c)$ tend to $q_0$, and
$P_t'(t)=3ct^2-4t+b$ tends to zero.  Away from its finitely many zeros,
the affine reconstruction \eqref{eq:affine-inverse} is defined and
$x=2/P_t'(t)$ is unbounded.  Hence $V(\Delta)\subseteq S_F$.

Conversely, take $q_0\notin V(\Delta)$ and a sufficiently small analytic
neighborhood $N$ on which the three projective roots remain distinct.  Every
finite root stays in a compact affine-root chart and has $P'$ bounded away
from zero; equations \eqref{eq:affine-inverse} therefore bound its source
branch.  A root can leave that chart only by tending to $[1:0]$, which forces
$c\to0$.  In the $s=V/U$ chart its equation and inverse are
\[
 c=2s-bs^2+2as^3,qquad
 x=\frac{s}{1-bs+3as^2},qquad y=b-3as,
\]
together with the polynomial expression \eqref{eq:projective-z}.  At
$s=c=0$, the derivative of the first equation with respect to $s$ is $2$.
The analytic implicit-function theorem therefore makes this branch regular,
and it tends to $(0,b,a-4b^2)$ rather than to infinity.  Thus every source
branch over $N$ is bounded, proving $q_0\notin S_F$.
\end{proof}

\begin{proposition}[normalization of the discriminant]\label{prop:disc-normalization}
The morphism
\[
 \nu:\A^2_{t,c}\longrightarrow V(\Delta),\qquad
 (t,c)\longmapsto(t^2-ct^3,\,4t-3ct^2,\,c)
\]
is the normalization.  It is birational and one-to-one over the triple-root
curve $\Gamma$.
\end{proposition}

\begin{proof}
The displayed point has a repeated root $t$, so it lies on $V(\Delta)$.
In the function field of $V(\Delta)$, $t$ satisfies the monic equation
$t^2-bt+3a=0$, hence is integral.  Away from $3bc=4$ it is recovered by
\[
 t=\frac{b-9ac}{4-3bc},
\]
so $\nu$ is birational.  Since $\A^2$ is normal and the finite integral
extension supplies all repeated-root parameters, it is the normalization.
On $\Gamma$ the unique triple root determines $t$, so the normalization has
one point over each point of $\Gamma$.
\end{proof}

The divisor $r=P'(t)=0$ in the polynomial graph coordinates
$(t,r,c)$ accounts for the lost length-two part of a generic discriminant
fiber.  This boundary statement concerns the finite marked-root cover, not
ramification of the etale affine map $F$.
