\section{The normalized incidence and reconstruction}\label{sec:weighted-incidence}

Retain all hypotheses and notation from Section~\ref{sec:weighted-map}.  For
a target $(A,B,C)$ define
\begin{equation}\label{eq:weighted-pencil}
 E_{A,B,C}(W)=H(W)-BCW+cAC^2.
\end{equation}

\subsection{Four distinct spaces}

The \emph{raw incidence} is
\begin{equation}\label{eq:raw-incidence}
 \mathcal I_H=V(E_{A,B,C}(W))
 \subset\A^3_{A,B,C}\times\A^1_W.
\end{equation}
If $h_n$ is the leading coefficient of $H$, then $h_n^{-1}E$ is monic in
$W$.  Hence $\mathcal I_H\to\A^3$ is finite flat of degree $n$, with basis
$1,W,\ldots,W^{n-1}$.  The projective homogenization has value $h_nW^n$ on
the hyperplane at infinity, so \eqref{eq:raw-incidence} already contains all
projective roots.

The polynomial $E$ is irreducible in $k[A,B,C,W]$.  Indeed, viewed as a
linear polynomial in $A$ over the UFD $k[B,C,W]$, it is primitive because
$cC^2$ and $H(W)-BCW$ are coprime; it is irreducible over the fraction field
and hence by Gauss's lemma.  Thus $\mathcal I_H$ is integral.  We do
\emph{not} assert that it is normal.

Let
\begin{equation}\label{eq:normalization}
 \nu:\widetilde{\mathcal I}_H\longrightarrow\mathcal I_H
\end{equation}
be its normalization.  This is the \emph{entire finite normal model} used
below.  Normality is imposed by taking normalization; it is not deduced for
the raw incidence.  Since a finite-type $k$-scheme is excellent,
\eqref{eq:normalization} is finite, and the composite
$\widetilde{\mathcal I}_H\to\A^3$ remains finite of degree $n$
\cite{stacks-normalization}.

The \emph{raw simple-root locus} is
\[
 \mathcal I_H^{\mathrm{simp}}=D(E_W)\subset\mathcal I_H,
 \qquad E_W=H'(W)-BC.
\]
Finally, define rational functions on the common function field by
\begin{equation}\label{eq:weighted-reconstruction}
 \gamma=\frac{BC-H'(W)}c,qquad
 x=\frac C\gamma,qquad
 y=\frac{W-\gamma}{C},qquad
 z=\frac{\gamma-1-a_0(W/\gamma-1)}{b_0x^2}.
\end{equation}
The \emph{regular-reconstruction open} is the explicitly defined open
\begin{equation}\label{eq:RH}
 \mathcal R_H=\left\{\xi\in\widetilde{\mathcal I}_H:
 x,y,z\text{ are regular at }\xi\right\}.
\end{equation}
It is open because the regularity domain of a rational function is open.
These four objects---raw incidence, normalization, simple-root locus, and
regular-reconstruction open---must not be identified without proof.

\begin{proposition}[reconstruction over \texorpdfstring{$C\ne0$}{C nonzero}]\label{prop:weighted-Cnonzero}
Over $D(C)$, normalization is unnecessary on the simple-root locus and
\[
 G_H^{-1}(D(C))\xrightarrow{\sim}
 \mathcal I_H^{\mathrm{simp}}\cap D(C).
\]
The inverse is given by \eqref{eq:weighted-reconstruction}.
\end{proposition}

\begin{proof}
On the incidence, $\gamma=-E_W/c$.  Thus on $D(C E_W)$ both $C$ and
$\gamma$ are units and all formulas \eqref{eq:weighted-reconstruction} are
regular.  Moreover $E_W$ is invertible there, so the raw incidence is etale
over the regular target on this locus; it is normal, and the normalization is
an isomorphism over it.  Put $u=W/\gamma$, $v=u-1$, and
$S=(\gamma-1-a_0v)/b_0$.  Then $x=C/\gamma$, $y=v/x$, and $z=S/x^2$.
Substitution in \eqref{eq:weighted-formulas}, using $E=0$, returns
$(A,B,C)$ and $W=u\gamma$.

Conversely, at a source point with $C=x\gamma\ne0$ one has
\[
 E_W(W(x,y,z))=-c\gamma\ne0.
\]
The same identities recover $(x,y,z)$.  They are equalities in localized
coordinate rings, not merely pointwise formulas.
\end{proof}

\begin{theorem}[global regular-reconstruction theorem]\label{thm:weighted-global}
For every admissible $H$ satisfying \eqref{eq:admissible-H}, the root marking
\[
 \iota_0:\A^3_{x,y,z}\longrightarrow\mathcal I_H,qquad
 (x,y,z)\longmapsto(G_H(x,y,z),W(x,y,z))
\]
lifts uniquely to an isomorphism
\[
 \iota:\A^3_{x,y,z}\xrightarrow{\sim}\mathcal R_H.
\]
Under this identification, $G_H$ is the restriction of the finite map
$\widetilde{\mathcal I}_H\to\A^3_{A,B,C}$.  Its generic degree is $n$.
\end{theorem}

\begin{proof}
The identity $E_{G_H(x,y,z)}(W(x,y,z))=0$ follows directly from
\eqref{eq:weighted-formulas} and defines $\iota_0$.  Proposition
\ref{prop:weighted-Cnonzero} makes this morphism dominant and birational.
Because $\A^3$ is normal, the universal property of normalization gives a
unique lift $\iota$ to $\widetilde{\mathcal I}_H$.

The lift is quasi-finite: each of its fibers is contained in a fiber of the
etale, hence quasi-finite, map $G_H$.  It is a quasi-finite birational
morphism to the normal scheme $\widetilde{\mathcal I}_H$, so Zariski's Main
Theorem makes it an open immersion \cite{stacks-zmt}.  On its image, the rational functions
\eqref{eq:weighted-reconstruction} pull back to the source coordinates;
hence the image lies in $\mathcal R_H$.

Conversely, wherever $x,y,z$ in \eqref{eq:weighted-reconstruction} are
regular, they define a morphism to $\A^3$.  The identities
$G_H(x,y,z)=(A,B,C)$ and $W=(1+xy)\gamma$, already proved in the common
function field on $D(C)$, hold on that regularity neighborhood.  This
morphism is inverse to the open immersion, so its domain is exactly
$\mathcal R_H$.

Finally, a generic target has $C\ne0$ and $E_W\ne0$ at each of the $n$ roots
of the monic degree-$n$ polynomial $E$.  Proposition
\ref{prop:weighted-Cnonzero} therefore reconstructs $n$ source points, proving
that the generic degree of $G_H$ is $n$.
\end{proof}

Because $\widetilde{\mathcal I}_H$ is normal, regularity of a rational
function is detected at height-one points.  Consequently the complement is
specified set-theoretically by
\begin{equation}\label{eq:boundary-valuations}
 \widetilde{\mathcal I}_H\setminus\mathcal R_H
 =\bigcup_{\substack{D\in\widetilde{\mathcal I}_H^{(1)}\\
 \min(v_D(x),v_D(y),v_D(z))<0}}
 \overline{\{D\}}.
\end{equation}
Thus the affine source is one explicitly defined open, and its divisorial
complement consists of the pole valuations \eqref{eq:boundary-valuations}.
The argument in this paper does not enumerate these valuations for every root
profile of $H$; complete boundary lists are stated only when the required
root hypotheses and local degree counts have been checked.

\subsection{The fiber over \texorpdfstring{$C=0$}{C=0}}

At $C=0$, the raw incidence specializes to $H(W)=0$.  The simple-root locus
is not the affine source globally: a repeated root at $W=0$ may have normalized
branches with regular reconstruction, while a simple additional root of $H$
may have a reconstruction pole.

The distinguished root $W=1$ is simple because $H'(1)=-c$.  In the completed
etale local incidence at that root,
\begin{equation}\label{eq:root-one-expansion}
 W=1-\frac BcC+
 \left(A+\frac{2+\kappa}{2c^2}B^2\right)C^2+O(C^3).
\end{equation}
The reconstruction then has $x\to0$ and
$y\to-(2+\kappa)B/c$; the choice \eqref{eq:a0} cancels the linear term in
$\gamma-1-a_0(W/\gamma-1)$, so formally $S=O(C^2)$ and $z=S/x^2$ has no
formal pole.
This completed-local calculation implies algebraic regularity: for the
excellent normal local ring of the normalized incidence, faithful flatness of
completion gives $K\cap\widehat{\mathcal O}=\mathcal O$ inside the completed
fraction field.

Write $H(W)=W^{m_0}L(W)$ with $m_0\ge2$ and $L(0)=h_0\ne0$.  The chart
$W=CR$, after division by $C^2$, has strict equation
\begin{equation}\label{eq:zero-cluster}
 C^{m_0-2}R^{m_0}L(CR)-BR+cA=0.
\end{equation}
For $m_0=2$, its generic special fiber is
$h_0R^2-BR+cA=0$; for $m_0\ge3$ and $B\ne0$, it has the generic solution
$R=cA/B$.  These formulas exhibit finite normalized branches but do not
classify every special point or every boundary divisor over $C=0$.
