\section{Examples and scope of the construction}\label{sec:examples}

\subsection{The cubic collision as three marked roots}

At the collision target $(-1/4,0,0)$, the binary cubic is
\[
 Q(U,V)=-2U^2V+\frac12V^3
       =-2V(U-V/2)(U+V/2).
\]
Its three simple projective roots are
\[
 [1:0],\qquad[1/2:1],\qquad[-1/2:1].
\]
Through the two inverse charts of Section~\ref{sec:cubic-incidence}, these
give respectively
\[
 (0,0,-1/4),\qquad(-1,3/2,13/2),\qquad(1,-3/2,13/2).
\]
This recovers the collision without solving the original three coordinate
equations simultaneously.

\subsection{Canonical admissible polynomials}

For every $d\ge2$, let
\[
 H_d(W)=W^d(1-W).
\]
Then $n=d+1$, $H_d(0)=H_d(1)=H_d'(0)=0$, and $H_d'(1)=-1$, so $c=1$.
Moreover $H_d''(1)=-2d$, hence $\kappa=-2d\ne-2$, and
\[
 a_0=\frac{1-2d}{2(d-1)}.
\]
For any $b_0\in k^\times$, equations
\eqref{eq:weighted-source-invariants}--\eqref{eq:weighted-formulas} therefore
give a degree-$(d+1)$ generic inverse cover with constant Jacobian $b_0$ and
monodromy $S_{d+1}$.  This is an application of the uniform theorems, not a
bounded computation.

At $C=0$, $H_d$ has a root of multiplicity $d$ at zero and a simple root at
one.  The only conclusions established here about this special fiber are those in
Section~\ref{sec:weighted-incidence}: the affine source is the
regular-reconstruction open of the normalized incidence, the root-one formal
chart is finite, and \eqref{eq:zero-cluster} gives the generic zero-cluster
chart.  No exact image or complete boundary classification for these examples
is asserted here.

\subsection{Logical scope}

The cubic source is literally the raw simple-projective-root locus.  The
weighted source is generally not: it is the regular-reconstruction open in
the normalization of the raw affine-root incidence.  The generic degree and
monodromy are uniform, while exact images and special-fiber boundary strata
depend on $H$ and lie outside the theorems proved in this paper.

\subsection{Comparison, verification, and provenance}

The explicit map \eqref{eq:F}, its determinant, and the rational collision
were publicly circulated in July 2026 in an announcement attributed in
contemporaneous sources to Levent Alp\"oge and Fable
\cite{alpoge-announcement,gallagher-explainer,zhang-consequences}.  The direct
announcement has not been archived in a form from which this paper can settle
the full discovery history, and we make no priority claim.

The organization by binary cubics with one simple projective root marked was
also described publicly by Andy Jiang \cite{jiang-marked-root}.  The present
paper supplies the two-chart scheme-theoretic proof, including the root at
infinity, and uses that formulation with attribution rather than as a claim
of independent discovery.  The weighted formulas in
Section~\ref{sec:weighted-map} overlap the same-day explanatory account
\cite{gallagher-explainer}; our contribution here is to state carefully the
raw incidence, its normalization, and the regular-reconstruction open.

Dean Cureton independently formalized the determinant, the three rational
evaluations, a determinant-one rescaling, and the complex specialization in
Lean 4 \cite{cureton-lean}.  The pinned implementation is available at
\href{https://github.com/deancureton/jacobian/commit/0d4a9212d874226ad81ce5a926becddfa94e6a88}{the audited commit}.
It does not formalize Theorems~\ref{thm:marked-root} or
\ref{thm:weighted-global}.  We link to that implementation rather than copy
it; the audited commit did not contain an explicit license file.

Exact computer-algebra scripts also check the displayed determinant, both
cubic chart compositions, and representative weighted identities.  Those
checks are useful guards against transcription errors.  They are not used as
proofs of the uniform statements in $H$ or $n$.
