\section{Further directions}\label{sec:further}

The results above stop with the uniform C01--C04 construction.  Two natural
questions remain.  First, for a fixed admissible $H$, one may try to classify
the pole valuations \eqref{eq:boundary-valuations} and thereby determine the
exact image, nonproperness set, and special fibers of $G_H$.  Second, one may
compare other one-variable inverse constructions with the weighted family
under polynomial source and target automorphisms.

Subsequent work in this repository answers the first degree-by-degree
comparison question beyond the scope of this paper.  For every generic fiber
degree $N\ge4$, there are at least $\tau(N-1)$ pairwise stably inequivalent
three-dimensional Keller maps, where $\tau$ is the positive-divisor counting
function.  The cancellation types are indexed by the
proper divisors $r\mid N-1$, with $m=(N-1)/r-1$, and are separated by their
boundary ramification and nilpotency indices; a split weighted seed supplies
one further reduced-boundary class.  The common normalized-root-cover
framework and the proof are recorded in
\texttt{MARKED\_ROOT\_KELLER\_MAPS.md}.

Neither direction is used in this paper.  In particular, computations for
special quartics, bounded-degree discriminants, finite fields, transfer
blocks, and cancellation families do not prove or strengthen any theorem
stated above.  They may suggest conjectures or supply test cases, but the
present argument depends only on the explicit cubic identities, the two
projective root charts, normalization and Zariski's Main Theorem, and the
purity argument for symmetric monodromy.

A useful next foundational step would be an independent algebraic-geometry
audit of Theorem~\ref{thm:weighted-global}, especially the passage from the
formal $C=0$ expansions to algebraic regularity and the precise divisorial
structure of the complement of $\mathcal R_H$.  Until that audit is supplied,
the paper deliberately states no uniform seed-by-seed boundary theorem.
