\ifdefined\HurwitzLLPaper
\else

\section{The marked Hessian-divisor quotient}

For a weighted primitive $H$, let $D_H$ be the reduced discriminant of
$H(W)-sW+t$ and let
\[
 \nu_H:\A^1_r\longrightarrow D_H,
 \qquad r\longmapsto(H'(r),rH'(r)-H(r))                 \tag{6.1}
\]
be its affine normalization.  Write
$R_H=k[H'(r),rH'(r)-H(r)]\subset k[r]$.

\begin{proposition}[weighted intrinsic-boundary reconstruction]
\label{prop:weighted-intrinsic-boundary}
Assume that zero is an exact double root of $H$, that every nonzero root is
simple, and that the weighted reconstruction formulas are used on their
maximal regular open.  In the map-intrinsic package
$\mathcal B(F_H)$ of the construction-independent stable-normalization
theorem, the complete set of
target images of boundary divisors is a pair
\[
 (Z_\Delta,Z_0).
\]
It is intrinsically ordered: $Z_\Delta$ is the unique member receiving a
ramified boundary prime, and $Z_0$ is the other member.  Removing $Z_0$
recovers, from the map rather than from the chosen pencil coordinate, the
finite normalization morphism
\[
 \widetilde D_H\times\mathbb G_m
 \longrightarrow D_H\times\mathbb G_m.                \tag{6.1a}
\]
Consequently this ordered finite morphism is transported by polynomial
left--right equivalence and acquires only an affine-space factor under
stabilization.
\end{proposition}

\begin{proof}
Write the weighted inverse equation as
\[
 E(W)=H(W)-BCW+cAC^2.
\]
On $C\ne0$, regular reconstruction is equivalent to $E_W\ne0$.  At the
generic point of the reduced discriminant exactly one inverse root is double.
Its normalized critical branch is outside the etale source open and has
$(e,f)=(2,1)$; the other $N-2$ roots are simple affine branches.  Thus the
local degree sum is $2+(N-2)=N$.

At the generic point of $C=0$, the exact double root at zero gives the
residue-degree-two affine cluster, and the normalized root at one gives a
residue-degree-one affine branch.  If $\rho\ne0,1$ is another simple root,
the reconstruction identity
\[
 y=(W-\gamma)/C,
 \qquad \gamma\longrightarrow-H'(\rho)/c
\]
has a pole unless $-H'(\rho)/c=\rho$.  In the exceptional cancellation of
that numerator, the numerator of the reconstruction formula for $z$ tends
to $\rho-1$, so the branch is still outside the source open.  Hence the
remaining $N-3$ branches are unramified boundary branches and the second
degree sum is $2+1+(N-3)=N$.

Off the discriminant and $C=0$, every inverse root is simple and all
reconstruction coordinates are regular.  The two degree sums therefore
exclude an additional height-one prime over either divisor.  They also
exclude a height-one prime over any other target divisor, since regular
reconstruction covers its generic point.  There is no complement component
supported only in higher codimension: after removing the listed divisors,
choose an affine neighborhood $V$ of its hypothetical generic point.  The
intersection of the affine opens $V$ and $\A^3\subset\overline X_{F_H}$ is
affine because the normalization is separated, and normal Hartogs extension
gives
\[
 \Gamma(V,\mathcal O)=
 \Gamma(V\cap\A^3,\mathcal O).
\]
The corresponding inclusion of affine schemes is therefore an isomorphism,
a contradiction.  This proves completeness.  The discriminant member is
uniquely characterized by its ramified boundary prime, so the ordering is
intrinsic.

Finally, on $C\ne0$ the target change
\[
 (A,B,C)\longmapsto(s,t,C)=(BC,cAC^2,C)
\]
identifies $Z_\Delta\setminus Z_0$ with
$D_H\times\mathbb G_m$, and its normalized critical incidence is
$\widetilde D_H\times\mathbb G_m$.  Because both vertices were selected
inside $\mathcal B(F_H)$, this identification proves (6.1a) is the explicit
form of an intrinsic finite stratum map.  The last assertion now follows
from the map-intrinsic and stabilization clauses of that theorem.
\end{proof}

\begin{proposition}[full Fitting divisor]
\label{prop:full-fitting}
There is an equality of ideal sheaves on the normalization
\[
 \operatorname{Fitt}_0\Omega_{\widetilde D_H/D_H}=(H''(r)). \tag{6.2}
\]
This ideal, including every root multiplicity, is preserved by polynomial
left--right equivalence and by adjoining identity variables.
\end{proposition}

\begin{proof}
The tangent-map core theorem identifies the critical restriction with
$\nu_H$ and gives
$\Omega_{k[r]/R_H}\simeq k[r]/(H'')\,dr$, proving (6.2).
Proposition~\ref{prop:weighted-intrinsic-boundary} identifies the displayed
finite normalization morphism inside the map-intrinsic package.  The
construction-independent stable normalization theorem therefore transports
its relative differential module and Fitting ideals.  Under
stabilization one has
\[
 \Omega_{(\widetilde D_H\times\A^q)/(D_H\times\A^q)}
 \simeq\operatorname{pr}_1^*\Omega_{\widetilde D_H/D_H}.
\]
Zeroth Fitting ideals commute with this flat base change.  Equality of the
ideal sheaves preserves their valuations at every height-one point, hence
preserves the multiplicities in (6.2), not only the reduced support.
\end{proof}

If the support of (6.2) contains two points, the units of
$k[r,\xi,\xi^{-1},z_1,\ldots,z_q]$ and subtraction of the two corresponding
principal-prime equations force the induced coordinate change to have the
form $r\mapsto ar+b$, $a\ne0$.  Proposition~\ref{prop:full-fitting} then says
that the full effective Hessian divisor is carried affinely, with its
coefficients.

\begin{lemma}[intrinsic zero and infinity marks]
\label{lem:weighted-boundary-marks}
Assume zero is the unique multiple root of $H$ and is an exact double root.
The closure of the normalized discriminant stratum over
$Z_\Delta\cap Z_0$ has the unique finite valuation center $r=0$.  Thus the
second target boundary intrinsically marks zero.  If the Fitting divisor has
at least two support points, stable left--right equivalence also preserves
the unique point at infinity of the projective normalization.
\end{lemma}

\begin{proof}
On the normalized discriminant stratum,
\[
 B=\frac{H'(r)}C,
 \qquad
 A=\frac{rH'(r)-H(r)}{cC^2}.                           \tag{6.2a}
\]
If $C\to0$ and $r\to\rho$ has finite target limit, then
$H'(\rho)=H(\rho)=0$.  The unique-multiple-root hypothesis forces
$\rho=0$.  Conversely, writing $H(r)=h_2r^2+O(r^3)$ and setting $r=Ck$
gives
\[
 B\to2h_2k,
 \qquad A\to(h_2/c)k^2,
 \qquad B^2=4ch_2A,                                   \tag{6.2b}
\]
so this center actually occurs.  Since it is defined as the valuation center
of the closure of the intrinsic finite stratum (6.1a), it is transported by
the map-intrinsic normalization theorem.

With two Fitting-support points, the preceding unit argument makes the
induced normalization-coordinate change affine.  It therefore extends to
the projective completions and fixes their unique points outside the affine
lines, proving the infinity assertion.
\end{proof}

On the boundary-clean seed locus with at least two Fitting-support points,
Lemma~\ref{lem:weighted-boundary-marks} canonically marks $r=0$ and
$r=\infty$.  The minimal
stable invariant used for generic moduli is therefore
\[
 \mathcal H_H=(\P^1;\operatorname{div}(H''),0,\infty).
\]
It remembers the full effective Hessian divisor, including multiplicities,
but does not remember node pairing or the conductor.  Write $d=N-2$ and let
\[
 \mathcal Q_d=[U_d/\mathbb G_m],\qquad r\longmapsto ar,
\]
where $U_d$ is the open of squarefree effective degree-$d$ divisors disjoint
from zero and infinity.  The assignment $H\mapsto\mathcal H_H$ defines the
marked Hessian-divisor map to $\mathcal Q_d$.

\begin{proposition}[nonempty Hessian-clean locus]
\label{prop:ordinary-boundary-clean}
For every $N\ge4$, the normalized admissible degree-$N$ seed space contains
a nonempty open on which $H''$ is squarefree and avoids zero and infinity,
zero is the unique multiple primitive root, and all other roots of $H$ are
simple.  On this locus $\mathcal H_H$ and the nonzero root sheets are defined
without node-pairing or conductor data.
\end{proposition}

\begin{proof}
The generic discriminant theorem proves that the squarefree-Hessian open
meets the normalized admissible slice in every degree.  The remaining
conditions are open.  Boundary-cleanness is nonempty on the normalized slice,
as witnessed by
\[
 H_N^{\mathrm{split}}(W)=\frac12W^2(1-W)(1+W^{N-3}).
\]
Indeed,
\[
 (H_N^{\mathrm{split}})'(1)=-1,\qquad
 (H_N^{\mathrm{split}})''(1)=-(N+1)\ne-2,\qquad
 (H_N^{\mathrm{split}})''(0)=1.
\]
Put $P_N=2H_N^{\mathrm{split}}/W^2=(1-W)(1+W^{N-3})$.
This polynomial is squarefree and has no zero root.  At every root $\alpha$
of $P_N$,
\[
 \left.\frac{2(H_N^{\mathrm{split}})'(W)}W\right|_{W=\alpha}
 =\alpha P_N'(\alpha)\ne0.
\]
Thus $\gcd(H_N^{\mathrm{split}},(H_N^{\mathrm{split}})')=W$.
The normalized seed space is irreducible, so its nonempty Hessian and
boundary-clean opens intersect.
\end{proof}

\begin{proposition}[marked Hessian-divisor moduli]
\label{prop:hessian-divisor-unramified}
For every $N\ge4$, the marked Hessian-divisor map is generically unramified.
After restriction to a nonempty open it is etale of exact degree $N-2$ onto
its image, which has dimension $N-3$.  Its geometric fibers are exactly the
rerootings at the nonzero simple roots of the seed.
\end{proposition}

\begin{proof}
The stack $\mathcal Q_d$ is smooth of dimension $d-1=N-3$.  A normalized
first-order seed deformation
$H+\varepsilon\dot H$ which is trivial in the quotient of these divisors
by normalization-line scaling satisfies
\[
 \dot H''(r)=\alpha H''(r)+\beta rH'''(r).             \tag{6.7}
\]
The first term is the scalar ambiguity in a defining section and the second
is the infinitesimal action of $r\mapsto(1+\varepsilon\beta)r$; these are all
infinitesimal automorphisms because zero and infinity are marked.  Twice
integrating and using
$\dot H(0)=\dot H'(0)=0$ gives
\[
 \dot H=\alpha H+\beta(rH'-2H).
\]
The tangent endpoint equations now give
\[
 0=\dot H(1)=-\beta,
 \qquad 0=\dot H'(1)=-\alpha,
\]
so $\dot H=0$.  Thus the marked Hessian-divisor map is unramified; on the
locus with trivial divisor stabilizer it is etale after shrinking.

For the degree statement, if $a$ is a nonzero simple root of $H$, put
\[
 \kappa_a=-\frac1{aH'(a)},\qquad G_a(w)=\kappa_aH(aw).
\]
Then $G_a$ is normalized and
\[
 E_{G_a}(w;s,t)=\kappa_a E_H\left(aw;
        \frac{s}{\kappa_a a},\frac{t}{\kappa_a}\right). \tag{6.8}
\]
Moreover $G_a''(w)=\kappa_a a^2H''(aw)$, so every rerooting has the same
marked Hessian divisor.  Conversely, equality in $\mathcal Q_d$ gives
$G''(r)=cH''(ar)$.  Twice integrating and using the double zero at the
origin gives $G(r)=cH(ar)/a^2$; then $G(1)=0$ forces $H(a)=0$, and
$G'(1)=-1$ forces $c/a^2=\kappa_a$.  Thus every fiber point is one of the
$N-2$ distinct generic rerootings.  This proves the exact degree and the
dimension claim using only $\mathcal H_H$.
\end{proof}

\begin{lemma}[the affine root sheet]
\label{lem:affine-root-sheet}
On the Hessian-clean open, the root-one component over $C=0$ is
the unique unramified component of the inverse incidence which meets the
regular-reconstruction open outside the zero cluster.  The double-zero locus
forms a residue-degree-two cluster rather than a simple root sheet, and every
component belonging to another simple root lies in the reconstruction
boundary.
\end{lemma}

\begin{proof}
The tangent-core identities give
\[
 \gamma=-E_W/c,\qquad x=C/\gamma,
 \qquad y=(W-\gamma)/C.                               \tag{6.9}
\]
The exact double root at zero supplies the separate residue-degree-two chart
described by the quadratic zero-cluster model.  Let $\rho\ne0$ be a simple
root of $H$.  Along its component over $C=0$ one has
\[
 \gamma\longrightarrow-H'(\rho)/c.
\]
Unless this limit equals $\rho$, the last expression in (6.9) has a pole.
If it does equal $\rho$, the leading $y$-pole cancels, but from
\[
 z=\frac{\gamma-1-a_0(W/\gamma-1)}{b_0x^2}            \tag{6.10}
\]
the numerator tends to $\rho-1$.  Thus this branch is regular only for
$\rho=1$.  At root one, $-H'(1)/c=1$, and the weighted reconstruction
formulas are regular.  Boundary-cleanness makes the remaining roots simple
and separates their branches, proving uniqueness.  The characterization uses
only ramification and membership in the distinguished affine open, so it is
intrinsic and is preserved by stabilization.
\end{proof}

\begin{theorem}[marked Hessian-divisor faithfulness]
\label{thm:affine-mark-faithful}
On the Hessian-clean open, the marked Hessian divisor together with the
affine root sheet recovers the normalized seed exactly.  In
particular, for seeds $H,G$ on this open,
\[
 F_H\sim_{\mathrm{stable\ LR}}F_G\quad\Longrightarrow\quad H=G.        \tag{6.11}
\]
\end{theorem}

\begin{proof}
Stable normalization functoriality identifies the marked Hessian divisors
and carries the regular-reconstruction open, hence the affine root sheet, to
its counterpart.  Proposition~\ref{prop:hessian-divisor-unramified} says that two
seeds in one geometric Hessian-divisor fiber are related by a unique generic
rerooting
\[
 G=G_a,\qquad G_a(w)=\kappa_aH(aw),
 \qquad \kappa_a=-\frac1{aH'(a)},                    \tag{6.12}
\]
at a nonzero simple root $a$ of $H$.  Under the incidence isomorphism (6.8),
the root-one sheet of $G_a$ maps to the root-$a$ sheet of $H$.  By
Lemma~\ref{lem:affine-root-sheet}, preservation of the affine sheet forces
$a=1$.  Then $\kappa_1=1$ because $H'(1)=-1$, so $G=H$.
\end{proof}

\begin{theorem}[marked moduli normalization]
\label{thm:marked-moduli-normalization}
After shrinking the Hessian-clean seed open, its affine-marked
Hessian-divisor map is an isomorphism onto the normalization of its image.
Under this
identification, forgetting the affine root sheet is finite etale of degree
$N-2$ onto the marked Hessian-divisor image.
\end{theorem}

\begin{proof}
The seed open is smooth and hence normal.  The marked map is unramified
because its composite with the Hessian-divisor map is, and
Theorem~\ref{thm:affine-mark-faithful} makes it generically one-to-one.
It is therefore birational and quasi-finite onto the normalization of its
image.  Zariski's Main Theorem makes it an open immersion; shrinking the
target image gives the asserted isomorphism.  Proposition
\ref{prop:hessian-divisor-unramified} identifies the forgetful morphism with the
$N-2$ rerooting cover and proves the final assertion.
\end{proof}

\section{The stronger scheme-theoretic decoration}

The preceding D1 and F2 arguments use only $\mathcal H_H$ and the affine
root sheet.  The node-pairing scheme and conductor are retained separately
as the stronger D2 invariant; neither is an input to generic dimension,
unramifiedness, rerooting degree, or affine-mark faithfulness.

The self-intersection of (6.1) is intrinsic.  In two normalization
coordinates put
\[
 D_s(r,u)=\frac{H'(r)-H'(u)}{r-u},
\]
\[
 D_t(r,u)=
 \frac{rH'(r)-H(r)-uH'(u)+H(u)}{r-u}.
\]
The genuine off-diagonal closure is the scheme
\[
 \Gamma_H^{\rm off}
 =V\bigl((D_s,D_t):(r-u)^\infty\bigr).                 \tag{6.13}
\]
The saturation removes the diagonal while retaining its isolated cusp
limits.  Its quotient by $(r,u)\leftrightarrow(u,r)$ records the two
normalization branches paired at each ordinary node.  Isomorphisms of
normalization morphisms preserve this construction, and stabilization takes
its product with affine space, so the node-pairing scheme is stable.

Let
\[
 \mathfrak c_H=\operatorname{Ann}_{\mathcal O_{D_H}}
 (\nu_{H*}\mathcal O_{\widetilde D_H}/\mathcal O_{D_H}) \tag{6.14}
\]
be the conductor.  On the ordinary cusp-node locus, if $P_H(r)$ cuts out the
normalization branches occurring in (6.13), then
\[
 \mathfrak c_Hk[r]=(H''(r)^2P_H(r)).                   \tag{6.14a}
\]
For arbitrary plane-curve singularities with primitive implicit equation
$f_H(S,T)=0$, plane-curve adjunction gives
\[
 \mathfrak c_Hk[r]=
 \left(
 \frac{(\partial f_H/\partial T)
 (H'(r),rH'(r)-H(r))}{H''(r)}
 \right).                                             \tag{6.14b}
\]
D2 retains the finite conductor morphism
\[
 \operatorname{Spec}(k[r]/\mathfrak c_Hk[r])
 \longrightarrow
 \operatorname{Spec}(R_H/(R_H\cap\mathfrak c_Hk[r])), \tag{6.14c}
\]
together with the full Fitting divisor and node pairing.  On the
boundary-clean locus it also retains the second-boundary center divisor,
whose pencil-coordinate presentation is $\gcd(H,H')$ but whose invariant
definition is the corresponding finite boundary incidence.  If the Fitting
support contains at least two points, the normalization-cylinder unit
argument forces its coordinate change to be affine and therefore preserves
the unique infinity point.  Without these intrinsic center and two-support
hypotheses, the center label or infinity point is not included merely because
it appears in a chosen pencil presentation.  With the stated hypotheses,
every retained part is preserved under stable polynomial left--right
equivalence.  The extended D2 audit records
the exact saturation, quotient, transversality, conductor, and collision
checks omitted from the minimal moduli proof.

\begin{remark}[coarse affine-mark descent]
The selected root extends tautologically on the marked admissible-cover
stack; the companion Hurwitz--LL paper records the resulting finite
representable etale quotient and its comparison with the repository covers
at arbitrary collisions.  The same paper proves the coarse descent: the
closure of the distinguished generic degree-one component in the contracted
finite root cover is finite birational over the normal marked coarse
compactification, hence isomorphic to it.  Thus the affine mark has one
coarse specialization over every DVR limit.  At a collision of multiplicity
$\mu$ the special fiber is $k[T]/(T^\mu)$: it retains the stacky length but
has one geometric point.  In particular, pair and triple collisions do not
give counterexamples.
\end{remark}

\fi

\ifdefined\DecoratedNormalizationPaper
\else

\begin{proposition}[Hurwitz--LL compactification]
\label{prop:hurwitz-ll-compactification}
For $P_{H,s}(W)=H(W)-sW$, the curve $D_H$ is the pullback of the
universal Lyashko--Looijenga point-in-branch-divisor incidence along
$s\mapsto P_{H,s}$.  The normalized seed locus has a proper compactification
$\overline{\mathcal S}^{\rm adm}_N$ obtained as the normalization of its
closure in the stack of marked admissible polynomial covers.  The generic
degree-$N-2$ rerooting relation extends to the proper marked/unmarked
groupoid.  On the collision-separating model it is a finite representable
etale quotient-stack morphism.  A total collision of $m$ zero-fiber points
has $m$-cycle monodromy on the special transverse slice
$\epsilon=\delta^m$, while generic coarse divisorial inertia is a
transposition.
\end{proposition}

\begin{proof}
A critical point $r$ of $P_{H,s}$ satisfies $s=H'(r)$, and its critical value
is
\[
 P_{H,s}(r)=H(r)-rH'(r)=-t.
\]
This proves the first assertion and recovers the normalization (6.1).
On the exact-double distinct-root open,
\[
 H(W)=-\frac{W^2(W-1)\prod_{i=1}^{N-3}(W-\rho_i)}
                  {\prod_{i=1}^{N-3}(1-\rho_i)},
\]
so the labelled source is the configuration space of the $N-3$ points
$\rho_i$ on the thrice-marked rational curve.  Its collision-separating
compactification is $\overline M_{0,N}$; quotienting by $S_{N-3}$ forgets
their labels.  Equivalently, the seed defines a polynomial Hurwitz cover
with total ramification over infinity, a marked ramified point $p_0$ over
zero, and a marked unramified point $p_1$ over zero.  Taking the normalization
of its closure in the proper admissible-cover stack gives
$\overline{\mathcal S}^{\rm adm}_N$.\footnote{For the normalization and
twisted-cover interpretation of the admissible-cover stack, see D. Abramovich,
A. Corti and A. Vistoli,
\href{https://arxiv.org/abs/math/0106211}{\emph{Twisted bundles and admissible
covers}}.  Collision-allowing intermediate models are constructed in
A. Deopurkar,
\href{https://arxiv.org/abs/1206.4535}{\emph{Compactifications of Hurwitz
spaces}}.}

Put $n=N-2$.  Forgetting which of the $n$ simple zero-fiber points is
designated $p_1$, while retaining all of them as an unordered divisor, is the
quotient morphism
\[
 [\overline M_{0,N}/S_{n-1}]
 \longrightarrow[\overline M_{0,N}/S_n].             \tag{6.15}
\]
It is finite, representable, and etale of degree
$[S_n:S_{n-1}]=n$, and restricts to the rerooting relation (6.8).  On the
cyclic slice $u^m=\epsilon$, normalization writes
$\epsilon=\delta^m$ and cycles the $m$ choices.  For $m>2$ this slice passes
through the codimension-$(m-1)$ total-collision locus; the generic coarse
discriminant instead has one pair collision and transposition inertia.
Finally, the stable-map branch divisor extends
in flat families, so the LL incidence extends over the closure.\footnote{B.
Fantechi and R. Pandharipande,
\href{https://arxiv.org/abs/math/9905104}{\emph{Stable maps and branch
divisors}}.}  Its finite part has degree $N-1$, after removing the fixed
branch contribution $(N-1)[\infty]$.
\end{proof}

\begin{proposition}[rerooting boundary pullback]
\label{prop:rerooting-boundary-pullback}
Let $\pi$ be (6.15).  For any boundary divisor whose distinguished component
contains exactly $k$ of the $n$ simple zero-fiber marks, write
$\Delta_{k,\mathrm{in}}$ and $\Delta_{k,\mathrm{out}}$ according as $p_1$
belongs to that component or not.  Then
\[
 \pi^*[\Delta_k]
  =[\Delta_{k,\mathrm{in}}]+[\Delta_{k,\mathrm{out}}], \tag{6.16}
\]
with restriction degrees $k$ and $n-k$.  After contraction to the coarse
coefficient collision model, the generic discriminant pullback consists of
one degree-one component of ramification index two and one unramified
component of degree $n-2$.
\end{proposition}

\begin{proof}
The quotient-stack map (6.15) is etale, so both coefficients in (6.16) are
one.  Above an unordered boundary configuration, selecting one of the $k$
points on the distinguished component gives the first restriction, while
selecting one of the remaining $n-k$ gives the second.  At a generic
coefficient collision, choosing the double value has local square-root
normalization and contributes $2\cdot1$ to the degree; choosing one of the
$n-2$ other values is unramified.  Thus the degree sum is
$2+(n-2)=n$.
\end{proof}

\begin{remark}
The number $N-2$ here is the subgroup index $[S_n:S_{n-1}]$, not the degree
of the classical LL morphism.  The latter has degree $N^{N-2}$ on the full
space of monic centered degree-$N$ polynomials.\footnote{See S. Lando and D.
Zvonkine,
\href{https://arxiv.org/abs/math/0303218}{\emph{Counting ramified coverings
and intersection theory on spaces of rational functions I}}, and the
references there for the classical polynomial LL degree.}  The fixed pencil and the
normalized seed locus have smaller dimensions, so an LL degree for their
restriction requires a separately specified Hurwitz stratum and intersection
problem.
\end{remark}

\begin{theorem}[marked-zero-fiber LL degree]
\label{thm:restricted-ll-degree}
Let $\mathcal A_N^\circ$ be the exact-double normalized seed locus on which
all critical values are simple.  Put $n=N-2$, let $v_1,\ldots,v_n$ be the
nonzero critical values, and write
\[
 \prod_{j=1}^n(V-v_j)=V^n+c_1V^{n-1}+\cdots+c_n.
\]
Then
\[
 \Lambda_N:\mathcal A_N^\circ\longrightarrow\mathbb P(1,2,\ldots,n),
 \qquad H\longmapsto[c_1:\cdots:c_n]
\]
is generically finite of degree
\[
 \boxed{\deg\Lambda_N=(N-2)N^{N-3}}.
\]
If $\overline{\mathcal A}^{\rm LL}_N$ is the normalized graph closure in the
marked admissible-cover compactification and
$\eta=c_1(\mathcal O_{\mathbb P(1,\ldots,n)}(1))$, then
\[
 \int_{\overline{\mathcal A}^{\rm LL}_N}
       \overline\Lambda_N^*\eta^{N-3}
 =\frac{N^{N-3}}{(N-3)!}.
\]
\end{theorem}

\begin{proof}
The classical polynomial LL map has degree $N^{N-2}$.  Generic
precomposition by $\mu_N$ acts freely on its monic centered source and
preserves every critical value.  Quotienting by affine source isomorphism
therefore leaves $N^{N-3}$ covers.  Requiring one specified critical value to
be zero is a transverse target slice and does not change that count.  Its
fiber profile is $(2,1^{N-2})$, so marking an unramified point over zero gives
$N-2$ choices.  The source conditions $p_0=0,p_1=1$ and the target condition
$H'(1)=-1$ then choose a unique normalized representative.  This proves the
degree formula.

The weighted-projective stack satisfies
$\int_{\mathbb P(1,\ldots,n)}\eta^{n-1}=1/n!$.  Multiplying by the generic
degree and using $n=N-2$ gives
\[
 \frac{(N-2)N^{N-3}}{(N-2)!}
 =\frac{N^{N-3}}{(N-3)!},
\]
as asserted.
\end{proof}

\begin{theorem}[caustic and Maxwell boundary classes]
\label{thm:caustic-maxwell-boundary}
Put $n=N-2$ and work on
$\overline{\mathfrak B}^{\rm root}_N=[\overline M_{0,n+2}/S_n]$.
Its boundary-orbit classes satisfy the unique relation
\[
 \sum_{k=1}^{n-1}k(n-k)\Delta_{0,k}
 =\sum_{k=2}^{n}k(k-1)\Delta_k^{\rm free}.           \tag{6.17}
\]
Thus
\[
 \mathcal B_n=\{\Delta_2^{\rm free},\ldots,
 \Delta_n^{\rm free},\Delta_{0,1},\ldots,\Delta_{0,n-2}\}
\]
is a basis of the rational Picard group.  If $\mathcal C_N$ and
$\mathcal M_N$ denote the closures of the caustic and Maxwell divisors, then
in the redundant orbit presentation
\[
\begin{aligned}
 [\mathcal C_N]
  &=-\sum_{k=2}^{n}(k-1)(k-2)\Delta_k^{\rm free}
    +\sum_{k=1}^{n-1}k(n-k)\Delta_{0,k},\\
 [\mathcal M_N]
  &=-\frac12\sum_{k=2}^{n}(k-1)(k-2)(k-3)
                     \Delta_k^{\rm free}\\
  &\qquad+\frac12\sum_{k=1}^{n-1}k(n-k)(n+k-2)\Delta_{0,k}.
\end{aligned}                                                   \tag{6.18}
\]
In the basis $\mathcal B_n$ these become
\[
 [\mathcal C_N]=2\sum_{k=2}^{n}(k-1)\Delta_k^{\rm free},       \tag{6.19}
\]
and
\[
\begin{aligned}
 [\mathcal M_N]
  ={}&\frac12\sum_{k=2}^{n}(k-1)
       \bigl((2n-3)k-(k-2)(k-3)\bigr)\Delta_k^{\rm free}\\
    &-\frac12\sum_{k=1}^{n-2}k(n-k)(n-k-1)\Delta_{0,k}.
\end{aligned}                                                   \tag{6.20}
\]
Moreover $[\mathcal C_N]=2\kappa_1$.
\end{theorem}

\begin{proof}
The logarithmic differential $dH/H$ has residue vector
$(2,1^n,-N)$.  Every proper partial sum is nonzero, so on a stable pointed
source it has a nonzero residue at every node and its zero divisor avoids the
nodes.  If $K$ is the relative dualizing class and $D$ the marked divisor,
the double-zero locus has class
\[
 \pi_*((K+D)(2K+D))
 =2\bigl(\sum_i\psi_i-\delta\bigr)=2\kappa_1.
\]
The symmetric boundary expansion of $\kappa_1$, together with the averaged
Keel relation (6.17), gives (6.18)--(6.19) for the caustic.

For Maxwell, label the simple roots temporarily, normalize one of them to
$1$, and write $H=W^2\prod_i(W-a_i)$ and $K_H=H'/W$.  On the interior,
\[
 \operatorname{Disc}_V\operatorname{Res}_W(K_H,V-H)
 =\text{unit}\cdot\mathcal C_N^3\mathcal M_N^2.
\]
A free collision of $k$ roots gives orders
$(k-1)(k-2)$, $(k-1)(k-2)(k-3)/2$, and $k(k-1)(k-2)$
for caustic, Maxwell, and the left side.  If $k$ roots tend to zero, the
first two orders are $k(k-1)$ and $k(k-1)^2/2$.  If $q=n-k$ roots tend to
infinity, their pole orders are
\[
 q(q-1+2k),\qquad
 \frac q2\bigl((n-1)(q-1+2k)+k(k-1)\bigr).
\]
Averaging over the $n$ choices of normalized simple root gives the
$\Delta_{0,k}$ coefficients in (6.18), while a free collision contributes
the negative of its vanishing order.  Finally (6.17) gives (6.20).
The powers $3$ and $2$ have positive pullback orientation.  On the stable
target its discriminant pullback also contains target-boundary components;
the pole terms above have merely been moved across the principal-divisor
equality, so (6.18) does not suppress that boundary contribution.
\end{proof}

\begin{theorem}[formal comparison at arbitrary root collisions]
\label{thm:admissible-formal-comparison}
Pull the marked admissible-cover compactification of
Proposition~\ref{prop:hurwitz-ll-compactification} back along the pencil
$H-sW$ and mark a point over $-t$.  The normalized Stein factor of this
incidence is the canonical finite cover
\[
 X_H=\operatorname{Spec}k[s,t,W]/(H(W)-sW+t),
\]
and the normalized Stein factor of its critical-point incidence is
$\widetilde D_H=\operatorname{Spec}k[r]$ with map (6.1).  After contraction
of all admissible source and target bubbles, their completed boundary charts
and normalization--conductor square agree with the repository constructions,
including at simultaneous collisions of arbitrarily many root clusters.

More explicitly, if the distinct multiple roots are $\rho_i$, of
multiplicity $m_i$, and
\[
 e_i=m_i-1,\qquad E=\sum_i e_i,
\]
then the conductor in the completed normalization over $(s,t)=(0,0)$ is
\[
 \overline{\mathfrak c}_{(0,0)}
 =\bigoplus_i u_i^{e_i(E-1)}k[[u_i]].                 \tag{6.21}
\]
\end{theorem}

\begin{proof}
On the dense nondegenerate locus the marked-point incidence is
$H(W)-sW+t=0$.  This hypersurface is smooth, since its $t$-derivative is one,
and is finite of degree $N$ over $k[s,t]$.  Its coordinate ring and the
normalized Stein algebra of the proper admissible incidence are finite normal
$k[s,t]$-algebras with the same generic function algebra.\footnote{We use
the Noetherian Stein factorization theorem in its relative-normalization form;
see \href{https://stacks.math.columbia.edu/tag/03H0}{Stacks Project, Tag
03H0}.}  By uniqueness of
integral closure they coincide.  The identical argument for the critical
marking gives $k[r]$ and the map
$r\mapsto(H'(r),rH'(r)-H(r))$.

At any primitive root $\alpha$, put $W=\alpha+u$.  Completion and elimination
of $t$ give
\[
 \widehat{\mathcal O}_{X_H,(0,0,\alpha)}=k[[s,u]],
 \qquad t=s(\alpha+u)-H(\alpha+u),                    \tag{6.22}
\]
so Hensel separation makes simultaneous distinct clusters a product of the
same regular charts on both sides.  The reconstruction coordinates are the
same rational functions in the common function field, so normality identifies
their pole divisors and maximal regularity opens as well.

For a multiple root $\rho_i$, write
$H(\rho_i+u)=c_i u^{m_i}+O(u^{m_i+1})$.  On its discriminant branch,
with $q_i=t-\rho_i s$, formula (6.1) gives
\[
 \operatorname{ord}_{u_i}s=e_i,\qquad
 \operatorname{ord}_{u_i}q_i=e_i+1.                  \tag{6.23}
\]
Thus its intrinsic conductor exponent is $e_i(e_i-1)$.  Distinct roots give
distinct tangent lines $t=\rho_i s$, hence
$I(D_i,D_j)=e_i e_j$.  Applying the adjunction conductor formula (6.14b) to
$f=\prod_i f_i$ adds these pairwise intersections on branch $i$ and yields
\[
 e_i(e_i-1)+\sum_{j\ne i}e_i e_j=e_i(E-1),
\]
which proves (6.21).  Since both constructions have the same finite
normalization map, quotienting by this common conductor gives the same
conductor square.  The argument descends from geometric branches by Galois
invariance.
\end{proof}

\begin{theorem}[stacky marked extension]
\label{thm:stacky-marked-extension}
Put $n=N-2$.  The selected unramified zero-fiber root extends as the
tautological marked point on the collision-separating admissible-cover stack,
and forgetting the selection is
\[
 [\overline M_{0,N}/S_{n-1}]
 \longrightarrow [\overline M_{0,N}/S_n].
\]
This morphism is finite, representable, and etale of degree $n$.  After
contracting the admissible source and target bubbles, its normalized Stein
factors are the canonical finite root cover and discriminant normalization,
and its completed root-cover charts and normalization--conductor squares
agree with the repository constructions at every simultaneous multicluster
collision.
\end{theorem}

\begin{proof}
The marked point is part of the data of the marked admissible cover, so it
persists over the entire proper stack.  Proposition
\ref{prop:hurwitz-ll-compactification} identifies the forgetful map with the
displayed subgroup quotient; its finiteness, representability, etaleness,
and degree are therefore the subgroup-index calculation
$[S_n:S_{n-1}]=n$.  The assertions after contraction are exactly Theorem
\ref{thm:admissible-formal-comparison}.  Hence no additional collision-chart
or conductor gluing remains at the stack level.
\end{proof}

\begin{theorem}[coarse affine-mark descent]
\label{thm:coarse-affine-mark-descent}
Let $\overline S_N$ and $\overline B_N$ be the coarse spaces of the marked
and unmarked collision-separating compactifications.  In the relative
Hilbert space of length-one subschemes of the contracted finite root cover,
let $Z$ be the closure of the point selected generically by the distinguished
regular-reconstruction component.  Then $Z$ is finite and separated, and
\[
 Z\xrightarrow{\sim}\overline S_N.
\]
Consequently the selected affine mark has exactly one coarse specialization
over every DVR limit, independently of the admissible-cover lift and
compatibly with base change.
\end{theorem}

\begin{proof}
The coarse spaces are normal: etale-locally their rings are invariant
subrings of normal collision charts under finite symmetric groups.  By
Theorem~\ref{thm:stacky-marked-extension}, the contracted ambient root cover
is finite and agrees in every completed chart with the repository cover.  The
relative Hilbert space of its length-one subschemes is canonically the cover
itself.  Hence $Z$ is finite over $\overline S_N$.  It is integral and
generically the graph of the distinguished mark, so it is birational.  A finite birational
morphism to a normal integral algebraic space is an isomorphism.

For the explicit collision chart put $R=k[x_1,\ldots,x_\mu]$, let
$e_j$ be the elementary symmetric functions, and set $T=x_1$.  If $e'_j$
is elementary symmetric in $x_2,\ldots,x_\mu$, then
$e_j=e'_j+T e'_{j-1}$.  This recursion and the monic equation for $T$ give
\[
 R^{S_{\mu-1}}
 \simeq R^{S_\mu}[T]/
 \bigl(T^\mu-e_1T^{\mu-1}+\cdots+(-1)^\mu e_\mu\bigr).
\]
Thus the coarse marked chart is the universal finite flat root incidence.
At a total collision its fiber is $k[T]/(T^\mu)$, which has one geometric
point.  The categorical property of the coarse moduli map makes the stack
mark factor uniquely through this chart, and separatedness gives uniqueness
for every DVR specialization.  Products of these charts handle simultaneous
clusters.
\end{proof}

\fi

\ifdefined\HurwitzLLPaper
\else

\begin{corollary}[degreewise weighted stable moduli]
\label{cor:weighted-stable-moduli}
For every $N\ge4$, weighted degree-$N$ maps contain an
$(N-3)$-dimensional family of stable polynomial left--right classes.
\end{corollary}

\begin{proof}
By Proposition~\ref{prop:ordinary-boundary-clean}, the domain of the marked
Hessian-divisor map contains a nonempty open in every degree.
The map is generically finite by the preceding argument and its invariant is
constant on stable classes.  Its stable-class image therefore has dimension
$N-3$.
\end{proof}

\fi
