\section{Stable functoriality}

The construction-independent repository source is
\href{../../verified/STABLE_NORMALIZATION_FUNCTORIALITY.md}
{Stable normalization functoriality}.  It also records the relative
differential, Fitting-ideal, fiber-product, and conductor base-change clauses.
The proposition below is the paper-local specialization needed for the
boundary signature.

\begin{proposition}[stable normalization functoriality]
\label{prop:stable-normalization}
Let $Y=\operatorname{Spec}A$ be an integral affine variety over $\C$, let
$L/k(Y)$ be a finite extension, and let $B$ be the integral closure of $A$
in $L$.  Put $X=\operatorname{Spec}B=\Norm_Y(L)$.  Let $U\subset X$ be a
dense open and $\partial=(X\setminus U)_{\mathrm{red}}$.  For $s\ge0$, write
\[
 Y_s=Y\times\A^s,\qquad U_s=U\times\A^s,
 \qquad L_s=L(t_1,\ldots,t_s).
\]
Then there is a canonical isomorphism over $Y_s$
\[
 \Norm_{Y\times\A^s}\bigl(L(t_1,\ldots,t_s)\bigr)
 \simeq X\times\A^s.                                    \tag{5.1}
\]
In particular, taking $L=k(U)$ and $X=\overline X_F$ gives
\[
 \Norm_{Y\times\A^s}\bigl(k(U)(t_1,\ldots,t_s)\bigr)
 \simeq\overline X_F\times\A^s.
\]
Under (5.1) the following statements hold.
\begin{enumerate}
\item The reduced boundary is
\[
 (X\times\A^s\setminus U_s)_{\mathrm{red}}
 =\partial\times\A^s,                                   \tag{5.2}
\]
and its prime divisors are exactly $E\times\A^s$, with $E$ a
prime divisor of $\partial$.
\item If a boundary prime $E$ lies over a target prime divisor $Z$ whose
generic local ring in $Y$ is a DVR---as it is for the normal targets used
below---the corresponding stabilized primes are $E_s=E\times\A^s$ and
$Z_s=Z\times\A^s$.  Their divisorial valuations are the Gauss extensions
\[
 v_{E_s}|_L=v_E,\qquad v_{E_s}(t_i)=0,
\]
and
\[
 e(E_s/Z_s)=e(E/Z),\qquad f(E_s/Z_s)=f(E/Z).             \tag{5.3}
\]
\item Every reduced or scheme-theoretic multiple intersection commutes with
stabilization.  In particular, for target divisors $Z_1,\ldots,Z_q$ and
boundary divisors $E_1,\ldots,E_p$,
\[
 \bigcap_i(Z_i\times\A^s)
 =\left(\bigcap_i Z_i\right)\times\A^s,
 \qquad
 \bigcap_j(E_j\times\A^s)
 =\left(\bigcap_j E_j\right)\times\A^s                 \tag{5.4}
\]
as schemes.
\item If $R$ is the affine coordinate ring of one of these intersections,
then the stabilized ring is $R[t_1,\ldots,t_s]$ and
\[
 \Nil(R[t_1,\ldots,t_s])=\Nil(R)[t_1,\ldots,t_s].       \tag{5.5}
\]
Consequently reducedness and the exact nilpotency index of the nilradical
are preserved.
\item Scheme-theoretic images of finite stratum maps commute with
stabilization.  Thus every target stratum reconstructed as such an image
acquires exactly an $\A^s$ factor.
\end{enumerate}
\end{proposition}

\begin{proof}
This is the specialization of the construction-independent
\href{../../verified/STABLE_NORMALIZATION_FUNCTORIALITY.md}
{stable-normalization functoriality theorem}.  Apply its normalization,
boundary, valuation, intersection, nilradical, and scheme-image clauses to
the pair $(X,U)$.
\end{proof}

\begin{proposition}[map-intrinsic normalization package]
\label{prop:map-intrinsic-normalization}
Let $F:U\to Y$ be dominant, separated, and quasi-finite, with $U$ and $Y$
normal integral affine varieties.  Put
\[
 \overline X_F=\Norm_Y k(U),\qquad
 \mathcal B(F)=
 \bigl(\overline X_F\xrightarrow{\pi_F}Y,
       j_F:U\hookrightarrow\overline X_F,
       (\overline X_F\setminus j_F(U))_{\rm red}\bigr). \tag{5.6}
\]
Then $\mathcal B(F)$ is reconstructed functorially from $F$.  More precisely,
every pair of isomorphisms $a:U\simeq U'$ and $b:Y\simeq Y'$ satisfying
$G\circ a=b\circ F$ induces a unique isomorphism
$\overline a:\overline X_F\simeq\overline X_G$ over $b$ satisfying
$\overline a\circ j_F=j_G\circ a$.  It transports the boundary primes,
their target scheme-images, valuation labels, and every finite stratum map
selected uniquely from those data.  Moreover
\[
 \mathcal B(F\times\mathrm{id}_{\A^s})
 \simeq \mathcal B(F)\times\A^s.                       \tag{5.7}
\]
Consequently every decoration constructed functorially from this package is
invariant under stable polynomial left--right equivalence.
\end{proposition}

\begin{proof}
Zariski's Main Theorem gives the canonical open immersion
$j_F:U\hookrightarrow\overline X_F$.  The compatible function-field
isomorphisms induced by $a$ and $b$ identify the two finite extensions and
their integral closures.  This gives $\overline a$, and uniqueness of the
factorization through the integral closure gives
$\overline a\circ j_F=j_G\circ a$.  Complements, boundary primes, target
scheme-images, valuations, and intrinsic selection rules are therefore
transported.  Proposition~\ref{prop:stable-normalization} identifies the
normalization after stabilization, and uniqueness of the Zariski--Main
factorization identifies its open immersion with
$j_F\times\mathrm{id}_{\A^s}$.  This proves (5.7).  Composing (5.7) with the
isomorphism supplied by a left--right equivalence of stabilizations proves
the last assertion.
\end{proof}

\begin{lemma}[stable boundary signature]\label{lem:stable}
For the maps in Lemma~\ref{lem:boundary}, the pair $(e_\Delta,\mu)$ of
(2.1) is invariant under polynomial left--right equivalence and after
adjoining any number of identity variables.
\end{lemma}

\begin{proof}
A left--right equivalence transports the map-intrinsic package of
Proposition~\ref{prop:map-intrinsic-normalization}.  It therefore identifies
the finite covers, carries the distinguished affine open to the
distinguished affine open, and preserves boundary valuations, their $e,f$
labels, and scheme-theoretic target intersections.  The intrinsic
ramification marking from Lemma~\ref{lem:boundary} prevents the two target
vertices from being exchanged.

For stabilization, Proposition~\ref{prop:stable-normalization} identifies the
normalization and the distinguished open pair, preserves all reduced
boundary primes and their valuation labels, and base-changes the marked
target intersection scheme.  Its final assertion preserves reducedness and
the exact nilpotency index.  Thus both entries of $(e_\Delta,\mu)$ are
unchanged.
\end{proof}
