\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsthm,mathtools,microtype}
\usepackage[hidelinks]{hyperref}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{remark}[theorem]{Remark}
\newcommand{\A}{\mathbb A}
\renewcommand{\P}{\mathbb P}
\newcommand{\C}{\mathbb C}

\title{Hurwitz--LL Compactification of Rerooting}
\author{Repository working draft}
\date{\today}

\def\HurwitzLLPaper{1}

\begin{document}
\maketitle

\begin{abstract}
We compactify the rerooting relation for normalized marked-root pencils by
passing to marked admissible polynomial covers.  The resulting quotient-stack
description extends the generic degree-$N-2$ rerooting map, identifies its
boundary pullbacks, and compares the normalized Stein factors and conductor
square with the affine incidence constructions at arbitrary simultaneous root
collisions.  We also compute the degree $(N-2)N^{N-3}$ of the marked-zero-
fiber Lyashko--Looijenga morphism and the caustic and Maxwell divisor classes
on the root-boundary compactification.
\end{abstract}

\section{Setup, provenance, and review status}

\begin{quote}
\textbf{Audit notice (23 July 2026).}
The standalone \href{dvr-marking-audit.tex}{DVR marking audit} isolates the
missing global comparison.  The subsequent
\href{global-comparison-obstruction.tex}{global comparison obstruction}
shows that its first proposed arrow is false for the stated compactifications:
in degree five the target branch cross-ratio has leading term $x^3/y^2$, so
the admissible model requires a normalized graph blowup.  The abstract
quotient-stack map, finite-cover valuative lemma, and local root-collision
rings remain valid.  Claims identifying the full root-stable and
branch-stable compactifications should be read with the corrected graph
correspondence in that note.

The subsequent
\href{../../extended-geometry/BRANCH_SCALE_FAN.md}{branch-scale fan
experiment} computes the degree-six three-cluster chart.  Its radial part is
the predicted weighted braid fan, while a triple leading-value resonance
retains a second-scale cross-ratio and forces a further nested refinement.
The stable-target refinement in that chart is the normalized pullback of the
Kapranov four-point blowup of $\mathbb P^2$ giving $\overline M_{0,5}$; the
admissible source expansion and saturated-node normalization agree on the
generic equal-scale face and its first diagonal refinement.  All thirteen
radial scale types have verified degree-six source-component and saturated
node-monoid data.  All pairwise and triple Maxwell source-node rings and
their radial intersections are also verified; standard admissible-cover
deformation theory supplies their ambient toroidal gluing.  An exact central
Hurwitz enumeration finds two ambient classes, while the labelled
source-root cross-ratio selects the polynomial class as a reduced branch.
Thus the labelled coarse graphs agree in this chart.  Stack inertia, arising
from normalized label-preserving cover automorphisms, is trivial on every
radial type and is one diagonal $\mu_2$ at each pairwise or triple Maxwell
target node.  The matching algebraic candidate is the iterated second-root
stack along the four Maxwell boundary divisors.  Its face maps and
codimension-two inertia are explicit, and smooth tame-stack reconstruction
identifies it with the labelled ACV graph on this chart.  The separate
$(S_2)^3\rtimes S_3$ quotient is explicit.  Gluing under the full root-label
action is canonical on the global labelled normalized graph closure.  The
stable-target graph on every collision type is the normalized wonderful
pullback of the branch-diagonal building set.  The companion
\href{../../extended-geometry/SOURCE_VERTEX_RIGIDITY.md}{source-vertex
rigidity theorem} proves that two target-flag fibers and one third-fiber
equation uniquely reconstruct each rational source-component map.  The
general radial source theorem constructs those components and their
saturated node partitions on every ordered first-scale stratum for arbitrary
cluster multiplicities.  Polynomial monodromy subforests determine the
source dual trees, component degrees, and node indices on every nested
simple-branch resonance stratum.  Since the fully marked admissible-cover
branch morphism is finite, finite normalization proves that the wonderful
graph is already the complete coarse polynomial source graph.  Corrected H2
and coarse H3 are therefore unconditional.  Monodromy centralizers also
give the full-chain radial formula: equal cluster multiplicities have
trivial inertia, whereas unequal multiplicities can retain a finite
subgroup.  They also compute generic simple-resonance inertia from anchors
and node partitions.  The companion
\href{../../extended-geometry/RECURSIVE_RESONANCE_ATLAS.md}{recursive
resonance atlas theorem} gives normalized-initial-form formulas for source
flag divisors in framed residue coordinates, contraction descent by affine
composition, and simultaneous centralizer characters.  Its selected-factor
log-\'etale comparison supplies global coverage, overlap descent, and
identification with the normalized admissible-cover stack.  Thus
H1-COARSE, H1-STACK, H2, and coarse H3 are complete.
\end{quote}

Let $H$ be a normalized degree-$N$ seed with an exact double root at zero and
otherwise simple zero-fiber points on the dense open under consideration.
The construction-independent
\href{../../verified/TANGENT_MAP_CORE.md}{tangent-map core} supplies the
common normal form
\[
 (W,s)\longmapsto\bigl(s,Ws-H(W)\bigr).
\]
For
\[
 P_{H,s}(W)=H(W)-sW,
\]
its critical restriction, a point $r$ has $s=H'(r)$ and critical value
$H(r)-rH'(r)$.  Thus the discriminant normalization is
\[
 r\longmapsto \bigl(H'(r),rH'(r)-H(r)\bigr).
\]
The $N-2$ nonzero simple roots give the generic rerootings; collision strata
are treated in the marked admissible-cover closure.

The generic affine marking is supplied by the unique unramified root sheet
meeting the regular-reconstruction open.  On the marked admissible-cover
graph the selected root extends tautologically: if
$\Gamma_N^{\rm lab}$ is the fully labelled normalized graph, the
marked-to-unmarked map
\[
 [\Gamma_N^{\rm lab}/S_{N-3}]
 \longrightarrow
 [\Gamma_N^{\rm lab}/S_{N-2}]
\]
is finite, representable, and etale of degree $N-2$.  The normalized Stein
contractions, completed root-cover charts, and conductor squares are also
identified here at arbitrary simultaneous multicluster collisions.  After
coarse contraction, the finite birational closure of the distinguished
regular-reconstruction component is an isomorphism over the normal marked
coarse compactification, so the mark has a unique point over every DVR limit.

The admissible-cover technology cited below from Abramovich--Corti--Vistoli
and Deopurkar supplies an appropriate proper compactification framework.  It
does not externally review the repository-specific restricted LL-degree,
boundary-pullback, Picard, caustic, or Maxwell calculations.  No external
review of those calculations is recorded, and the statements below should be
read with that provenance distinction explicit.

\input{../marked-root-multiplicity/decorated-normalization.tex}

\end{document}
